---
output: github_document
---
<!--
Copyright 2023 Province of British Columbia

Licensed under the Apache License, Version 2.0 (the "License");
you may not use this file except in compliance with the License.
You may obtain a copy of the License at

http://www.apache.org/licenses/LICENSE-2.0

Unless required by applicable law or agreed to in writing, software distributed under the License is distributed on an "AS IS" BASIS,
WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
See the License for the specific language governing permissions and limitations under the License.
-->


<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# pahwq

<!-- badges: start -->
[![img](https://img.shields.io/badge/Lifecycle-Maturing-007EC6)](https://github.com/bcgov/repomountie/blob/master/doc/lifecycle-badges.md)
[![R-CMD-check](https://github.com/bcgov/pahwq/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/bcgov/pahwq/actions/workflows/R-CMD-check.yaml)
[![Codecov test coverage](https://codecov.io/gh/bcgov/pahwq/branch/main/graph/badge.svg)](https://app.codecov.io/gh/bcgov/pahwq?branch=main)
<!-- badges: end -->

## Overview

Implementation of the Photoxic Lipid Model (PTLM) for the calculation of Canadian
Water Quality Guidelines for Polycyclic Aromatic Hydrocarbons (PAH).

This package uses the Tropospheric Ultraviolet and Visible (TUV) Radiation Model (https://github.com/NCAR/TUV)
to calculate the light penetration through water of a given depth at a given location, with a specified Dissolved Organic Carbon concentration. The light exposure is then used (along with the PAH-specific molar absorption across a range of wavelengths), to calculate the light absorption (Pabs) of the given PAH at that location. This is then used to determine the phototoxic benchmark.

## Installation

### Windows

On Windows, you need to install [Rtools](https://cran.r-project.org/bin/windows/Rtools/).
Make sure that you install the appropriate version for your version of R
(i.e. Rtools 4.0 for R 4.0.x, Rtools 4.3 for R 4.3.x, etc.). This will install
a compiler toolchain on your computer (specifically `gfortran`) which is necessary
to compile the Fortran code for the TUV model.

### Mac

On a Mac, the easiest way to get started is to use [Homebrew](https://brew.sh/),
and install `gcc`, which includes `gfortran`:

```
brew install gcc
```

### Install pahwq

Once you have Rtools (Windows) or `gcc` (Mac), you can install the package with:

``` r
# install.packages("pak")
pak::pak("bcgov/pahwq")
```

*Note: You may get some warnings when you install the package - these are from
the Fortran compiler because the Fortran code is quite old and does not fully
conform to modern standards. These warnings do not affect the package
functionality or accuracy.*

## Example usage

To calculate the acute phototoxic water quality guideline (benchmark) for
Anthracene at 0.25 m depth in Okanagan Lake on June 21, 2023, with a measured
DOC of 5 g/m^3, you can use the following code:

1. Load the pahwq package:

```{r}
library(pahwq)
```

2. Run the TUV model at your location to determine the incident light intensity
across the wavelength spectrum, at each time stamp within the time window specified:

```{r}
irrad <- tuv(
  depth_m = 0.25,
  lat = 49.601632,
  lon = -119.605862,
  DOC = 5,
  date = "2023-06-21",
  tzone = -8L
)
head(irrad)
```

3. Calculate the value of P~abs~ for the PAH of interest, using the results of the TUV model run:

```{r}
(Pabs <- p_abs(irrad, "Anthracene"))
```

4. Finally, calculate phototoxic benchmark in ðg/L, supplying the $P_{abs}$ value.

```{r}
phototoxic_benchmark(Pabs, pah = "Anthracene")
```

We can compare the phototoxic benchmark to the narcotic benchmark to see the effect of the photoxicity of the PAH:

```{r}
narcotic_benchmark("Anthracene")
```

### Options

pahwq creates a directory on your computer to store the TUV model input and output files. By default, the location of this is set automatically to a standard location (determined by `tools::R_user_dir`). You can change the location of this directory by setting the `pahwq.tuv_data_dir` option:

```{r}
#| eval: false
options("pahwq.tuv_data_dir" = "path/to/my/tuv/data")
```

### Running the TUV model step-by-step

Step 2 above can be broken down if you desire to inspect each phase of running
the TUV model:

1. Set the options for the TUV model run:

```{r}
set_tuv_aq_params(
  depth_m = 0.25,
  lat = 49.601632,
  lon = -119.605862,
  DOC = 5,
  date = "2023-06-21",
  tzone = -8L
)
```

2. Run the TUV model

```{R}
run_tuv()
```

3. Get the results of the TUV model run, as a data.frame of incident irradiation
at each timestamp and wavelength, at the given water depth at your location:

```{r}
irrad <- get_tuv_results(file = "out_irrad_y")
head(irrad)
```

### Calculating photoxicity for multiple chemicals

If you want to calculate a phototoxic benchmark for several chemicals at a particular site, you
can use the `pb_multi()` function.

First, you must calculate the absorption with `tuv()`:

```{r}
irrad <- tuv(
  depth_m = 0.25,
  lat = 49.601632,
  lon = -119.605862,
  DOC = 5,
  date = "2023-06-21",
  tzone = -8L,
  quiet = TRUE
)
```

Then, use the results from the TUV model run and a vector of chemical names:

```{r}
pb_multi(irrad, pahs = c("Anthracene", "Benzo(a)pyrene", "Fluorene"))
```


### Sensitivity to Kd, depth, and time of year

Use the function `sens_kd_depth()` to investigate the sensitivity of a given PAH
at a location to light attenuation (`DOC` or `Kd`) and depth over a set of dates.

There is also a plotting function, `plot_sens_kd_depth()` to make a static or
interactive heat map of the analysis.

```{r}
# varying DOC
out <- sens_kd_depth(
  "Anthracene",
  lat = 52,
  lon = -113,
  DOC = 3:5,
  depth_m = c(0.25, 0.5),
  date = c("2023-07-01", "2023-08-01")
)
head(out)

plot_sens_kd_depth(out, interactive = FALSE)
```

```{r}
# Varying Kd_ref
out2 <- sens_kd_depth(
  "benzo(a)pyrene",
  lat = 57,
  lon = -120,
  Kd_ref = seq(10, 50, by = 10),
  depth_m = c(0.25, 0.5, 0.75, 1),
  date = c("2023-05-01", "2023-07-01", "2023-09-01")
)

plot_sens_kd_depth(out2, interactive = FALSE)
```

### Calculating Pabs and phototoxic benchmarks with experimental data

If you have experimentally obtained data with a measured irradiance across a
range of wavelengths, you can calculate Pabs using `p_abs_single()` on a 2-column
data.frame. The first column must contain wavelengths and be called `wl`, and the second column contains irradiance values. Irradiance values must be in units of `"uW / cm^2 / nm"` (default) or `"W / m^2 / nm"`.

```r
irrad_df <- read.csv("my-irrad-data.csv")

head(irrad_df)
#>    wl          i
#> 1 280 0.05500000
#> 2 281 0.05000000
#> 3 282 0.04666667
#> 4 283 0.04000000
#> 5 284 0.06000000
#> 6 285 0.05500000

p_abs_single(irrad_df, pah = "anthracene")
#> [1] 0.027156
```

To simulate this exposure for a period of time (say 16 hours of daylight over
a two-day period), use the `time_multiplier` argument to get the Pabs for the
full exposure time. E.g., if the exposure was one second, and you need a 16h exposure, the multiplier would be 3600 s in 1h * 16 h (= 57600).

```r
p_abs_single(irrad_df, pah = "anthracene", time_multiplier = 3600 * 16)
#> [1] 1564.17
```

You can then use this value to calculate phototoxic benchmark with the `phototoxic_benchmark()` function. You
can either supply a narcotic benchmark value directly, or allow the function to calculate
the narcotic benchmark for the chemical supplied:


```r
# Calculate the Pabs for the simulated exposure:
p_abs_exp <- p_abs_single(irrad_df, pah = "anthracene", time_multiplier = 3600 * 16)

# Calculate the phototoxic benchmark:
phototoxic_benchmark(p_abs_exp, narc_bench = 58.4)
#> [1] 1.272051
```

## Sensitivity analysis

An analysis was conducted to test the sensitivity to various parameters in the
TUV model. The analysis can be viewed at: https://bcgov.github.io/pahwq/articles/sensitivity.html.

## Getting Help or Reporting an Issue

To report bugs/issues/feature requests, please file an [issue](https://github.com/bcgov/pahwq/issues/).

## How to Contribute

If you would like to contribute to the package, please see our
[CONTRIBUTING](CONTRIBUTING.md) guidelines.

Please note that this project is released with a [Contributor Code of Conduct](CODE_OF_CONDUCT.md). By participating in this project you agree to abide by its terms.

## License

```
Copyright 2023 Province of British Columbia

Licensed under the Apache License, Version 2.0 (the &quot;License&quot;);
you may not use this file except in compliance with the License.
You may obtain a copy of the License at

http://www.apache.org/licenses/LICENSE-2.0

Unless required by applicable law or agreed to in writing, software distributed under the License is distributed on an &quot;AS IS&quot; BASIS,
WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
See the License for the specific language governing permissions and limitations under the License.
```
