---
title: "Sensitivity Analysis"
bibliography: references.bib
editor_options:
  chunk_output_type: inline
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 8,
  fig.height = 8
)

local({
  hook_old <- knitr::knit_hooks$get("source")  # save the old hook
  knitr::knit_hooks$set(source = function(x, options) {
    if (!is.null(options$hide_src) && options$hide_src) {
      return(
        c("<details>
          <summary style='font-size:0.8em; margin-bottom:0; color: #666;'>
          Show/Hide Code
          </summary>\n",
          hook_old(x, options),
          "</details>")
      )
    }
    hook_old(x, options)
  })
})
```

```{r}
#| include: false
library(sf)
library(mapview)
library(rems)
library(dplyr)
library(tidyr)
library(gt)
library(ggplot2)
library(ggrepel)
library(ggtext)

theme_set(theme_bw())
theme_update(
  plot.title = ggtext::element_textbox_simple(margin = ggplot2::margin(10,0,10,0)),
  plot.caption = ggtext::element_textbox_simple(margin = ggplot2::margin(10,0,10,0)),
  axis.title.x = ggtext::element_textbox_simple(halign = 0.5)
)

sites_sf <- structure(list(emsid = c("0400390", "0500236", "E207466"), name = c("CHARLIE L DEEP STATION 1.2 KM EAST OF PARK",
"OKANAGAN L D/S KELOWNA STP (DEEP)", "QUAMICHAN LAKE; CENTRE"
), lon = c(-120.9642, -119.5134, -123.6625), lat = c(56.3125,
49.8614, 48.8003), elev_m = c(693, 342, 25), date = structure(c(1629823800,
1631127600, 1629918000), tzone = "Etc/GMT+8", class = c("POSIXct",
"POSIXt")), DOC = c(14, 4.26, 6.61), doc_min = c(0.96, 4.06,
6.37), doc_max = c(15.4, 5.17, 11.8), geometry = structure(list(
    structure(c(-120.9642, 56.3125), class = c("XY", "POINT",
    "sfg")), structure(c(-119.5134, 49.8614), class = c("XY",
    "POINT", "sfg")), structure(c(-123.6625, 48.8003), class = c("XY",
    "POINT", "sfg"))), class = c("sfc_POINT", "sfc"), precision = 0, crs = structure(list(
    input = "EPSG:4326", wkt = "GEOGCRS[\"WGS 84\",\n    ENSEMBLE[\"World Geodetic System 1984 ensemble\",\n        MEMBER[\"World Geodetic System 1984 (Transit)\"],\n        MEMBER[\"World Geodetic System 1984 (G730)\"],\n        MEMBER[\"World Geodetic System 1984 (G873)\"],\n        MEMBER[\"World Geodetic System 1984 (G1150)\"],\n        MEMBER[\"World Geodetic System 1984 (G1674)\"],\n        MEMBER[\"World Geodetic System 1984 (G1762)\"],\n        MEMBER[\"World Geodetic System 1984 (G2139)\"],\n        ELLIPSOID[\"WGS 84\",6378137,298.257223563,\n            LENGTHUNIT[\"metre\",1]],\n        ENSEMBLEACCURACY[2.0]],\n    PRIMEM[\"Greenwich\",0,\n        ANGLEUNIT[\"degree\",0.0174532925199433]],\n    CS[ellipsoidal,2],\n        AXIS[\"geodetic latitude (Lat)\",north,\n            ORDER[1],\n            ANGLEUNIT[\"degree\",0.0174532925199433]],\n        AXIS[\"geodetic longitude (Lon)\",east,\n            ORDER[2],\n            ANGLEUNIT[\"degree\",0.0174532925199433]],\n    USAGE[\n        SCOPE[\"Horizontal component of 3D system.\"],\n        AREA[\"World.\"],\n        BBOX[-90,-180,90,180]],\n    ID[\"EPSG\",4326]]"), class = "crs"), n_empty = 0L, bbox = structure(c(xmin = -123.6625,
ymin = 48.8003, xmax = -119.5134, ymax = 56.3125), class = "bbox"))), row.names = c(NA,
-3L), sf_column = "geometry", agr = structure(c(emsid = NA_integer_,
name = NA_integer_, lon = NA_integer_, lat = NA_integer_, elev_m = NA_integer_,
date = NA_integer_, DOC = NA_integer_, doc_min = NA_integer_,
doc_max = NA_integer_), levels = c("constant", "aggregate", "identity"
), class = "factor"), class = c("sf", "tbl_df", "tbl", "data.frame"
))

`%||%` <- function(x, y) if (is.null(x)) y else x
```

This report is to document the sensitivity of the TUV model, and the subsequent
calculation of water quality guidelines for PAHs, to variation in certain TUV
parameters.

## Input parameters

The input parameters for the calculation of light attenuation through water for
the TUV model are:

  * Vertical Attenuation Coefficient at wavelength `wvl`: $kvat = k_de^{-Sk(ref\_wvl-wvl)}$
      - **`Kd`** Light attenuation coefficient (set directly or calculated from DOC)
      - **`Sk`** (default `0.018`)
      - **`ref_wvl`** Reference wavelength for `Kd` (default `305`)
  * **`depth_m`** Water depth in m
  * **`lat`** Latitude in decimal degrees
  * **`lon`** Longitude in decimal degrees
  * **`elev_m`** Surface elevation, m above sea level
  * **`year`** Year
  * **`month`** Month
  * **`day`** Day
  * **`tzone`** Timezone  Local Time - UTC (default `0`; UTC)
  * **`tstart`** Start time, hours local time (default `0`)
  * **`tstop`** Stop time, hours local time (default `23`)
  * **`tsteps`** Number of time steps (default `24`)
  * **`albedo`** Surface albedo (default `0.07`)
  * **`o3_tc`** Ozone column, Dobson Units (DU) - Looked up in climatology (recommended default `300`)
  * **`so2_tc`** SO2 column, DU (default `0`)
  * **`no2_tc`** NO2 column, DU (default `0`)
  * **`taucld`** Cloud optical depth (default `0`)
  * **`zbase`** Cloud base, km (default `4`)
  * **`ztop`** Cloud top, km (default `5`)
  * **`tauaer`** Aerosol optical depth at 550 nm - Looked up in climatology (recommended default `0.235`)
  * **`ssaaer`** Aerosol single scattering albedo (default `0.990`)
  * **`alpha`** Aerosol Angstrom exponent(default `1.0`)
  * **`wvl_start`** Starting wavelength, nm (default `279.5`)
  * **`wvl_end`** End wavelength, nm (default `420.5`)
  * **`wvl_steps`** Number of wavelength intervals (default `141`)
  * **`nstr`** TUV run type; use -2 for fast, 4 for slightly more accurate (default `-2`)

Those currently under consideration for sensitivity analysis are:

  * **`Kd`** Light attenuation coefficient - includes use of **`DOC`** for calculating `Kd`
  * **`albedo`** Surface albedo (default `0.07`)
  * **`o3_tc`** Ozone column, Dobson Units (DU) - Looked up in climatology (recommended default `300`)
  * **`tauaer`** Aerosol optical depth at 550 nm - Looked up in climatology (recommended default `0.235`)
  * **`ssaaer`** Aerosol single scattering albedo (default `0.990`)
  * **`lat`** Latitude in decimal degrees
  * **`elev_m`** Surface elevation, m above sea level
  * **`depth_m`** Water depth in m
  * **`nstr`** TUV run type; use -2 for fast, 4 for slightly more accurate (default `-2`)

The analyses are conducted at three different sample lakes (Southern Interior,
Vancouver Island, Northeast), using two PAHs (Anthracene and Benzo[a]pyrene).

For each analysis, we will test a range of reasonable values for the parameter of
interest, and plot the phototoxic benchmark value calculated across that range. To look at the
relative effect of that parameter on the photoxicity of the PAH, we also plot
the ratio of phototoxic:narcotic benchmark.

Initially, sensitivity analyses will be univariate (varying the input of interest
while holding the others constant). If there are significant interactions expected
between certain variables, these can be explored.

## Basic usage of the `pahwq` package

To start, we will demonstrate a typical straightforward use case at a single site.

To calculate the acute phototoxic water quality guideline (phototoxic benchmark) for Anthracene
at 0.25 m depth in Okanagan Lake on June 21, 2023, with a measured DOC of 5 g/m^3,
you would use the following code:

First, set up the options for the model run:
```{r}
library(pahwq)

set_tuv_aq_params(
  depth_m = 0.25,
  lat = 49.601632,
  lon = -119.605862,
  elev_m = 342,
  DOC = 5,
  date = "2023-06-21",
  tzone = -8,
  albedo = 0.05
)
```

After setting them, you can view the options that will be used by TUV. Some are set via the function
inputs, some are looked up (e.g., `o3_tc`, `tauaer`) or calculated (e.g., `kd(305)` is calculated from the input `DOC`).

```{r}
view_tuv_aq_params()
```

Run the TUV model and get the results. The results show the underwater irradiance
at the specified depth, at each wavelength and time step.

```{r}
run_tuv()

# Get the results
tuv_res <- get_tuv_results(file = "out_irrad_y")
head(tuv_res)
```

We can inspect and verify the inputs that were used in the model run:

```{r}
tuv_run_params(tuv_res)
```

Next, calculate the site-specific light absorption ($P_{abs}$) for Anthracene from the
TUV results. The `p_abs()` function uses a lookup table to get the molar
absorption coefficient table for the specified PAH.

```{r}
(Pabs <- p_abs(tuv_res, "Anthracene"))
```

Finally, calculate phototoxic benchmark in ðg/L, supplying the $P_{abs}$ value.

```{r}
phototoxic_benchmark(Pabs, pah = "Anthracene")
```

We can compare the phototoxic benchmark to the narcotic benchmark to see the effect of the photoxicity of the PAH:

```{r}
narcotic_benchmark("Anthracene")
```

### A shortcut:

If you don't need to inspect every step of the way, the above process can be completed in two function calls:

```{r}
tuv_res <- tuv(
  depth_m = 0.25,
  lat = 49.601632,
  lon = -119.605862,
  elev_m = 342,
  DOC = 5,
  date = "2023-06-21",
  tzone = -8,
  albedo = 0.05
)

phototoxic_benchmark(tuv_res, "Anthracene")
```

## Sensitivity Analysis: Setup

We'll start by loading the packages we need and getting some data from B.C. EMS:

```{r setup, hide_src = TRUE}
#| eval: false
library(sf)
library(mapview)
library(rems)
library(dplyr)
library(ggplot2)
library(ggrepel)

con <- connect_historic_db()
db <- attach_historic_data(con)
sites_sf <- db |>
  filter(
    EMS_ID %in% c("0500236", "E207466", "0400390"),
    # grepl("(CHARLIE L)|(OKANAGAN)|(QUAMICHAN)", MONITORING_LOCATION),
    COLLECTION_START > as.Date("2020-01-01"), PARAMETER_CODE == "1103"
  ) |>
  collect() |>
  st_as_sf(coords = c("LONGITUDE", "LATITUDE"), crs = 4326, remove = FALSE) |>
  filter(!is.na(RESULT)) |>
  arrange(desc(COLLECTION_START)) |>
  group_by(EMS_ID, MONITORING_LOCATION, LONGITUDE, LATITUDE) |>
  mutate(doc_min = min(RESULT), doc_max = max(RESULT)) |>
  slice(1) |>
  select(emsid = EMS_ID,
         name = MONITORING_LOCATION,
         lon = LONGITUDE,
         lat = LATITUDE,
         date = COLLECTION_START,
         DOC = RESULT,
         doc_min,
         doc_max) |>
  left_join(
    tribble(
      ~ emsid, ~elev_m,
      "E207466", 25,
      "0400390", 693,
      "0500236", 342
    ), by = join_by("emsid")
  ) |>
  ungroup() |>
  relocate(elev_m, .after = lat)

disconnect_historic_db(con)
```


## Locations
```{r hide_src=TRUE}
mapview(sites_sf, zcol = "name", legend = FALSE)
```

### Okanagan Lake, Okanagan

* EMS ID: 0500236
* Region: Okanagan
* Latitude: 49.8614 N
* Longitude: 119.5134 W
* Site Depth: 70.1 m
* Maximum Lake Depth: 232 m
* Lake Elevation: 342 m
* Lake Surface Area: 350.08 sq km

### Charlie Lake, Peace

* EMS ID: 0400390
* Region: Peace
* Latitude: 56.3125 N
* Longitude: 120.9642 W
* Site Depth: 13 m
* Maximum Lake Depth: 13 m
* Lake Elevation: 693 m
* Lake Surface Area: 17.56 sq km

### Quamichan Lake, Vancouver Island

* EMS ID: E207466
* Region: Vancouver Island
* Latitude: 48.8003 N
* Longitude: 123.6625 W
* Site Depth: 7 m
* Maximum Lake Depth: 8 m
* Lake Elevation: 25 m
* Lake Surface Area: 2.88 sq km

### Prepare the data for comparisons, including setting the date to be the same
for all sites:

```{r hide_src=TRUE}
sites <- sites_sf |>
  st_drop_geometry() |>
  mutate(date = "2023-08-01")
gt(sites)
```

## Basic analysis using defaults

To perform the sensitivity analysis, we need to define a function, `multi_pb()` that allows us to do repeated runs of the TUV model and phototoxic benchmark calculation using a range of inputs at multiple sites:

```{r multi-tuv, hide_src=TRUE}
multi_pb <- function(df, site = "name", pah, varying, vals = NULL, ...) {
  if (!varying %in% union(names(tuv_aq_defaults()), names(formals(set_tuv_aq_params)))) {
    stop(varying, " is not a valid argument for `set_tuv_aq_params()`")
  }

  if (!is.null(vals)) {
    var_df <- data.frame(vals)
    names(var_df) <- varying

    df <- df |>
      select(!any_of(varying)) |>
      dplyr::cross_join(var_df)
  }

  df |>
    rowwise() |>
    mutate(
      tuv_res = list(calc_tuv(
          date = date,
          lat = lat,
          lon = lon,
          elev_m = elev_m,
          varying = .data[[varying]],
          vary_var = varying,
          ...
        ))
    ) |>
    cross_join(data.frame(PAH = pah)) |>
    mutate(
      timing = attr(tuv_res, "timing"),
      Pabs = p_abs(tuv_res, PAH),
      phototoxic_benchmark = phototoxic_benchmark(Pabs, pah = PAH),
      narcotic_benchmark = narcotic_benchmark(PAH),
      p_n_ratio = phototoxic_benchmark / narcotic_benchmark
    ) |>
    ungroup()
}

calc_tuv <- function(date, lat, lon, elev_m, varying, vary_var, ...) {
  args <- c(
    varying,
    list(
      depth_m = 0.25,
      date = as.Date(date),
      lat = lat,
      lon = lon,
      elev_m = elev_m
    ),
    ...
  )
  names(args)[1] <- vary_var

  # allow overriding of one of the core args by one supplied in 'varying',
  # this will keep the first of duplicated argument names, which will
  # be the one in 'varying'
  args <- args[unique(names(args))]

  do.call("set_tuv_aq_params", args)

  t <- system.time(run_tuv(quiet = TRUE))
  res <- get_tuv_results(file = "out_irrad_y")
  attr(res, "timing") <- unname(t["elapsed"])
  res
}
```

We then use the `multi_pb()` function to calculate $P_{abs}$, phototoxic benchmark, and ratio
of phototoxic:narcotic benchmark for Anthracene using recorded DOC values at the three sites.
This also uses the utility of pahwq to look up ozone column and aerosol optical
depth from climatologies based on latitude, longitude, and month.

```{r hide_src=TRUE}
diff_DOC <- multi_pb(sites, pah = c("Anthracene", "Naphthalene"), varying = "DOC")
```

```{r}
#| echo: false
diff_DOC |>
  select(-doc_min, -doc_max, -tuv_res) |>
  gt() |>
  data_color(columns = phototoxic_benchmark, palette = "Greens")
```

To compare the sites using the same input parameters other than location (lat, lon, elevation), we modify the data to set a constant [DOC] (DOC = 5), and calculate $P_{abs}$, phototoxic benchmark, and ratio of phototoxic:narcotic benchmark for Anthracene using constant DOC = 5:

```{r hide_src=TRUE}
same_DOC <- sites |>
  mutate(DOC = 5) |>
  multi_pb(pah = "Anthracene", varying = "DOC")
```

```{r}
#| echo: false
same_DOC |>
  select(-doc_min, -doc_max, -tuv_res) |>
  gt() |>
  data_color(columns = phototoxic_benchmark, palette = "Greens")
```

## Light Attenuation Coefficicent (_k~d~_)

The TUV model calculates the light attenuation coefficient $k_d(Î»)$ for each
wavelength, based on a reference $k_d(Î»_{ref})$, where $\lambda_{ref}$ is the reference wavelength.

$k_d(Î»_{ref})$ at 305nm ($k_{d,305}$) can be estimated from Dissolved Organic
Carbon concentration [DOC] using the equation:

$k_{d,305} = a_{305}[DOC]^{b_{305}} + 0.13$;
$a_{305}$ = 2.76 and $b_{305}$ = 1.2 from @morris1995; Equation 4-1a in @aris2023

For the sensitivity analysis for $k_d$, it is more interpretable to supply
a range of $k_d$ values to the TUV model directly, even though in practice it is
more likely to use [DOC] to estimate $k_d$.

We will test the sensitivity to $k_d$ using sample values from Table 3 (@aris2023), from 0.08 to 275:

```{r hide_src=TRUE}
kd_vals <- seq(0.08, 275, length.out = 20)
```

### Effect of varying _k~d~_ on phototoxic benchmark and ratio of phototoxic:narcotic benchmark for Anthracene

```{r hide_src=TRUE}
kd_test_a <- multi_pb(sites,
                       pah = "Anthracene",
                       varying = "Kd_ref",
                       vals = kd_vals,
                       Kd_wvl = 305,
                       o3_tc = 300,
                       tauaer = 0.235)

ggplot(kd_test_a, aes(x = Kd_ref, y = phototoxic_benchmark)) +
  geom_point() +
  facet_wrap(vars(name), ncol = 1) +
  labs(
    title = "Phototoxic benchmark for Anthracene at 3 sites in B.C. for a range of k~d~(305)",
    x = "k~d~(305)",
    y = "Phototoxic benchmark (Î¼g/L)"
  )

ggplot(kd_test_a, aes(x = Kd_ref, y = p_n_ratio)) +
  geom_point() +
  facet_wrap(vars(name), ncol = 1) +
  labs(
    title = "Ratio of phototoxic:narcotic benchmark for Anthracene at 3 sites in B.C. for a range of k~d~(305)",
    x = "k~d~(305)",
    y = "Ratio of phototoxic:narcotic benchmark"
  )
```

For Anthracene, the ratio of phototoxic:narcotic benchmark approaches 1 when $k_d(305)$ is ~
250 --- in other words at that level of $k_d$, phototoxic benchmark is nearly the same as narcotic benchmark,
indicating that the light is attenuated to such an extent that the phototoxicity
of Anthracene is not activated at the light levels when $k_d(305)$ >= 250.

According to the above equation for estimating $k_d$ from [DOC]:

$250 = 2.76[DOC]^{1.23} + 0.13$

[DOC] ~= 40

Therefore $k_d(305)$ of ~250 is representative of light attenuation when [DOC]
is ~40. This is a significantly higher DOC concentration than observed in our
three sample lakes, and is outside the range for which the above equation is
recommended (it is recommended for [DOC] between 0.2 and 23 mg/L).

### Effect of varying _k~d~_ on phototoxic benchmark and ratio of phototoxic:narcotic benchmark for Benzo[a]pyrene

We will use the same input values as we did for Anthracene:

```{r hide_src=TRUE}
kd_test_b <- multi_pb(sites,
                       pah = "Benzo[a]pyrene",
                       varying = "Kd_ref",
                       vals = kd_vals,
                       Kd_wvl = 305,
                       o3_tc = 300,
                       tauaer = 0.235)

ggplot(kd_test_b, aes(x = Kd_ref, y = phototoxic_benchmark)) +
  geom_point() +
  facet_wrap(vars(name), ncol = 1) +
  labs(
    title = "Phototoxic benchmark for Benzo[a]pyrene at 3 sites in B.C. for a range of k~d~(305)",
    x = "k~d~(305)",
    y = "Phototoxic benchmark (Î¼g/L)"
  )

ggplot(kd_test_b, aes(x = Kd_ref, y = p_n_ratio)) +
  geom_point() +
  facet_wrap(vars(name), ncol = 1) +
  labs(
    title = "Ratio of phototoxic:narcotic benchmark for Benzo[a]pyrene at 3 sites in B.C. for a range of k~d~(305)",
    x = "k~d~(305)",
    y = "Ratio of phototoxic:narcotic benchmark"
  )
```

## Surface Albedo (`albedo`)

We test surface albedo values from 0.05 to 0.1 (typical for water), and include
the suggested default of 0.7:

```{r}
albedo_vals <- c(seq(0.05, 0.1, length.out = 10), 0.07)
```

For this and the remaining analyses, we will set DOC = 5 for all sites to ensure
a constant $k_d$ value. We can then test the range of surface albedo values at
all three sites for Anthracene and Benzo[a]pyrene.

### Effect of varying Surface Albedo on phototoxic benchmark and ratio of phototoxic:narcotic benchmark for Anthracene

```{r hide_src=TRUE}
albedo_test_a <- multi_pb(sites,
                           pah = "Anthracene",
                           varying = "albedo",
                           vals = albedo_vals,
                           DOC = 5,
                           o3_tc = 300,
                           tauaer = 0.235)

ggplot(albedo_test_a, aes(x = albedo, y = phototoxic_benchmark)) +
  geom_point() +
  facet_wrap(vars(name), ncol = 1) +
  geom_point(
    data = filter(albedo_test_a, albedo == 0.07),
    colour = "red"
  ) +
  geom_text_repel(
    data = filter(albedo_test_a, albedo == 0.07, name == "QUAMICHAN LAKE; CENTRE"),
    colour = "red",
    label = "Recommended default = 0.07",
    nudge_x = 0.001,
    nudge_y = 0.05
  ) +
  labs(
    title = "Phototoxic benchmark for Anthracene at 3 sites in B.C. for a range of Surface Albedo",
    x = "Surface Albedo",
    y = "Phototoxic benchmark (Î¼g/L)"
  )

ggplot(albedo_test_a, aes(x = albedo, y = p_n_ratio)) +
  geom_point() +
  facet_wrap(vars(name), ncol = 1) +
  geom_point(
    data = filter(albedo_test_a, albedo == 0.07),
    colour = "red"
  ) +
  geom_text_repel(
    data = filter(albedo_test_a, albedo == 0.07, name == "QUAMICHAN LAKE; CENTRE"),
    colour = "red",
    label = "Recommended default = 0.07",
    nudge_x = 0.001,
    nudge_y = 0.0005
  ) +
  labs(
    title = "Ratio of phototoxic:narcotic benchmark for Anthracene at 3 sites in B.C. for a range of\nSurface Albedo",
    x = "Surface Albedo",
    y = "Ratio of phototoxic:narcotic benchmark"
  )
```

### Effect of varying Surface Albedo on phototoxic benchmark and ratio of phototoxic:narcotic benchmark for Benzo[a]pyrene

```{r hide_src=TRUE}
albedo_test_b <- multi_pb(sites,
                           pah = "Benzo[a]pyrene",
                           varying = "albedo",
                           vals = albedo_vals,
                           DOC = 5,
                           o3_tc = 300,
                           tauaer = 0.235)

ggplot(albedo_test_b, aes(x = albedo, y = phototoxic_benchmark)) +
  geom_point() +
  facet_wrap(vars(name), ncol = 1) +
  geom_point(
    data = filter(albedo_test_b, albedo == 0.07),
    colour = "red"
  ) +
  geom_text_repel(
    data = filter(albedo_test_b, albedo == 0.07, name == "QUAMICHAN LAKE; CENTRE"),
    colour = "red",
    label = "Recommended default = 0.07",
    nudge_x = 0.001,
    nudge_y = 0.0005
  ) +
  labs(
    title = "Phototoxic benchmark for Benzo[a]pyrene at 3 sites in B.C. for a range of Surface Albedo",
    x = "Surface Albedo",
    y = "Phototoxic benchmark (Î¼g/L)"
  )

ggplot(albedo_test_b, aes(x = albedo, y = p_n_ratio)) +
  geom_point() +
  facet_wrap(vars(name), ncol = 1) +
  geom_point(
    data = filter(albedo_test_b, albedo == 0.07),
    colour = "red"
  ) +
  geom_text_repel(
    data = filter(albedo_test_b, albedo == 0.07, name == "QUAMICHAN LAKE; CENTRE"),
    colour = "red",
    label = "Recommended default = 0.07",
    nudge_x = 0.001,
    nudge_y = 0.0003
  ) +
  labs(
    title = "Ratio of phototoxic:narcotic benchmark for Benzo[a]pyrene at 3 sites in B.C. for a range of\nSurface Albedo",
    x = "Surface Albedo",
    y = "Ratio of phototoxic:narcotic benchmark"
  )
```

## Ozone Column (`o3_tc`)

The pahwq package currently uses average monthly ozone column data from 1980-1991
from @fortuin1998, which is bundled with the TUV model. Based on
latitude and longitude and month, the ozone column value is looked up and supplied
to the TUV model. This default behaviour can be overridden by supplying a value to
the `o3_tc` argument in the `set_tuv_aq_params` function. In the absence of
climatological data, the recommended default value is 300 DU (@aris2023).

To test the sensitivity of phototoxic benchmark to ozone column, a range of values is tested from
280-420 DU, which are typical values in middle to Northern latitudes. See [NASA's 'Ozone Watch' page](https://ozonewatch.gsfc.nasa.gov/facts/dobson_SH.html) for more information on atmospheric ozone.

```{r hide_src=TRUE}
ozone_vals <- seq(280, 420, by = 10)
```

### Effect of varying Ozone Total Column on phototoxic benchmark and ratio of phototoxic:narcotic benchmark for Anthracene

```{r hide_src=TRUE}
ozone_test_a <- multi_pb(sites,
                          pah = "Anthracene",
                          varying = "o3_tc",
                          vals = ozone_vals,
                          DOC = 5,
                          tauaer = 0.235)

ggplot(ozone_test_a, aes(x = o3_tc, y = phototoxic_benchmark)) +
  geom_point() +
  facet_wrap(vars(name), ncol = 1) +
  geom_point(
    data = filter(ozone_test_a, o3_tc == 300),
    colour = "red"
  ) +
  geom_text_repel(
    data = filter(ozone_test_a, o3_tc == 300, name == "QUAMICHAN LAKE; CENTRE"),
    colour = "red",
    label = "Recommended default = 300",
    nudge_x = 5,
    nudge_y = 0.03
  ) +
  labs(
    title = "Phototoxic benchmark for Anthracene at 3 sites in B.C. for a range of Ozone Total Column",
    x = "Ozone Total Column",
    y = "Phototoxic benchmark (Î¼g/L)"
  )

ggplot(ozone_test_a, aes(x = o3_tc, y = p_n_ratio)) +
  geom_point() +
  facet_wrap(vars(name), ncol = 1) +
  geom_point(
    data = filter(ozone_test_a, o3_tc == 300),
    colour = "red"
  ) +
  geom_text_repel(
    data = filter(ozone_test_a, o3_tc == 300, name == "QUAMICHAN LAKE; CENTRE"),
    colour = "red",
    label = "Recommended default = 300",
    nudge_x = 5,
    nudge_y = 0.0005
  ) +
  labs(
    title = "Ratio of phototoxic:narcotic benchmark for Anthracene at 3 sites in B.C. for a range of\nOzone Total Column",
    x = "Ozone Total Column",
    y = "Ratio of phototoxic:narcotic benchmark"
  )
```

### Effect of varying Ozone Total Column on phototoxic benchmark and ratio of phototoxic:narcotic benchmark for Benzo[a]pyrene

```{r hide_src=TRUE}
ozone_test_b <- multi_pb(sites,
                          pah = "Benzo[a]pyrene",
                          varying = "o3_tc",
                          vals = ozone_vals,
                          DOC = 5,
                          tauaer = 0.235)

ggplot(ozone_test_b, aes(x = o3_tc, y = phototoxic_benchmark)) +
  geom_point() +
  facet_wrap(vars(name), ncol = 1) +
  geom_point(
    data = filter(ozone_test_b, o3_tc == 300),
    colour = "red"
  ) +
  geom_text_repel(
    data = filter(ozone_test_b, o3_tc == 300, name == "QUAMICHAN LAKE; CENTRE"),
    colour = "red",
    label = "Recommended default = 300",
    nudge_x = 5,
    nudge_y = 0.0005
  ) +
  labs(
    title = "Phototoxic benchmark for Benzo[a]pyrene at 3 sites in B.C. for a range of Ozone Total Column",
    x = "Ozone Total Column",
    y = "Phototoxic benchmark (Î¼g/L)"
  )

ggplot(ozone_test_b, aes(x = o3_tc, y = p_n_ratio)) +
  geom_point() +
  facet_wrap(vars(name), ncol = 1) +
  geom_point(
    data = filter(ozone_test_b, o3_tc == 300),
    colour = "red"
  ) +
  geom_text_repel(
    data = filter(ozone_test_b, o3_tc == 300, name == "QUAMICHAN LAKE; CENTRE"),
    colour = "red",
    label = "Recommended default = 300",
    nudge_x = 5,
    nudge_y = 0.0003
  ) +
  labs(
    title = "Ratio of phototoxic:narcotic benchmark for Benzo[a]pyrene at 3 sites in B.C. for a range of\nOzone Total Column",
    x = "Ozone Total Column",
    y = "Ratio of phototoxic:narcotic benchmark"
  )
```

## Aerosol Optical Depth (`tauaer`)

The pahwq package currently uses average monthly aerosol optical depth data from
2002 to 2023, obtained from MODIS/Aqua satellite data. Based on latitude and
longitude and month, the AOD value is looked up and supplied to the TUV model.
This default behaviour can be overridden by supplying a value to the `tauaer`
argument in the `set_tuv_aq_params` function. In the absence of
climatological data, the recommended default value is 0.235 (@aris2023).

To test the sensitivity of phototoxic benchmark to AOD, a range of values is tested from 0.1
(clear skies with maximum visibility) to 1.0 (very hazy skies). See the
[NASA Earth Observatory page on aerosols](https://earthobservatory.nasa.gov/global-maps/MODAL2_M_AER_OD).

```{r hide_src=TRUE}
aod_vals <- c(seq(0.1, 1.0, length.out = 10), 0.235)
```

### Effect of Varying Aerosol Optical Depth (AOD) on phototoxic benchmark and ratio of phototoxic:narcotic benchmark for Anthracene

```{r hide_src=TRUE}
aod_test_a <- multi_pb(sites,
                        pah = "Anthracene",
                        varying = "tauaer",
                        vals = aod_vals,
                        DOC = 5,
                        o3_tc = 300)

ggplot(aod_test_a, aes(x = tauaer, y = phototoxic_benchmark)) +
  geom_point() +
  facet_wrap(vars(name), ncol = 1) +
  geom_point(
    data = filter(aod_test_a, tauaer == 0.235),
    colour = "red"
  ) +
  geom_text_repel(
    data = filter(aod_test_a, tauaer == 0.235, name == "QUAMICHAN LAKE; CENTRE"),
    colour = "red",
    label = "Recommended default = 0.235",
    nudge_y = 0.1
  ) +
  labs(
    title = "Phototoxic benchmark for Anthracene at 3 sites in B.C. for a range of Aeorosol Optical Depth",
    x = "Aeorosol Optical Depth",
    y = "Phototoxic benchmark (Î¼g/L)"
  )

ggplot(aod_test_a, aes(x = tauaer, y = p_n_ratio)) +
  geom_point() +
  facet_wrap(vars(name), ncol = 1) +
  geom_point(
    data = filter(aod_test_a, tauaer == 0.235),
    colour = "red"
  ) +
  geom_text_repel(
    data = filter(aod_test_a, tauaer == 0.235, name == "QUAMICHAN LAKE; CENTRE"),
    colour = "red",
    label = "Recommended default = 0.235",
    nudge_y = 0.002
  ) +
  labs(
    title = "Ratio of phototoxic:narcotic benchmark for Anthracene at 3 sites in B.C. for a range of\nAeorosol Optical Depth",
    x = "Aeorosol Optical Depth",
    y = "Ratio of phototoxic:narcotic benchmark"
  )
```

### Effect of varying Aerosol Optical Depth (AOD) on phototoxic benchmark and ratio of phototoxic:narcotic benchmark for Benzo[a]pyrene

```{r hide_src=TRUE}
aod_test_b <- multi_pb(sites,
                        pah = "Benzo[a]pyrene",
                        varying = "tauaer",
                        vals = aod_vals,
                        DOC = 5,
                        o3_tc = 300)

ggplot(aod_test_b, aes(x = tauaer, y = phototoxic_benchmark)) +
  geom_point() +
  facet_wrap(vars(name), ncol = 1) +
  geom_point(
    data = filter(aod_test_b, tauaer == 0.235),
    colour = "red"
  ) +
  geom_text_repel(
    data = filter(aod_test_b, tauaer == 0.235, name == "QUAMICHAN LAKE; CENTRE"),
    colour = "red",
    label = "Recommended default = 0.235",
    nudge_y = 0.002
  ) +
  labs(
    title = "Phototoxic benchmark for Benzo[a]pyrene at 3 sites in B.C. for a range of\nAeorosol Optical Depth",
    x = "Aeorosol Optical Depth",
    y = "Phototoxic benchmark (Î¼g/L)"
  )

ggplot(aod_test_b, aes(x = tauaer, y = p_n_ratio)) +
  geom_point() +
  facet_wrap(vars(name), ncol = 1) +
  geom_point(
    data = filter(aod_test_b, tauaer == 0.235),
    colour = "red"
  ) +
  geom_text_repel(
    data = filter(aod_test_b, tauaer == 0.235, name == "QUAMICHAN LAKE; CENTRE"),
    colour = "red",
    label = "Recommended default = 0.235",
    nudge_y = 0.001
  ) +
  labs(
    title = "Ratio of phototoxic:narcotic benchmark for Benzo[a]pyrene at 3 sites in B.C. for a range of\nAeorosol Optical Depth",
    x = "Aeorosol Optical Depth",
    y = "Ratio of phototoxic:narcotic benchmark"
  )
```

## Aerosol Single Scattering Albedo (`ssaaer`)

Aerosol Single Scattering Albedo is currently set as a default constant value of
0.99. This can be overridden by setting the `ssaaer` parameter of the
`set_tuv_aq_params()` function to a different value.

Here we test a range for values from 0.80 to 0.99, including the recommended
default of 0.99 (@aris2023).

```{r hide_src=TRUE}
ssa_vals <- seq(0.8, 0.99, length.out = 10)
```

### Effect of varying Aerosol Single Scattering Albedo on phototoxic benchmark and ratio of phototoxic:narcotic benchmark for Anthracene

```{r hide_src=TRUE}
ssa_test_a <- multi_pb(sites,
                        pah = "Anthracene",
                        varying = "ssaaer",
                        vals = ssa_vals,
                        DOC = 5,
                        o3_tc = 300,
                        tauaer = 0.235)

ggplot(ssa_test_a, aes(x = ssaaer, y = phototoxic_benchmark)) +
  geom_point() +
  facet_wrap(vars(name), ncol = 1) +
  geom_point(
    data = filter(ssa_test_a, ssaaer == 0.99),
    colour = "red"
  ) +
  geom_text_repel(
    data = filter(ssa_test_a, ssaaer == 0.99, name == "QUAMICHAN LAKE; CENTRE"),
    colour = "red",
    label = "Recommended default = 0.99",
    nudge_y = .06
  ) +
  labs(
    title = "Phototoxic benchmark for Anthracene at 3 sites in B.C. for a range of\nAeorosol Single Scattering Albedo",
    x = "Aeorosol Single Scattering Albedo",
    y = "Phototoxic benchmark (Î¼g/L)"
  )

ggplot(ssa_test_a, aes(x = ssaaer, y = p_n_ratio)) +
  geom_point() +
  facet_wrap(vars(name), ncol = 1) +
  geom_point(
    data = filter(ssa_test_a, ssaaer == 0.99),
    colour = "red"
  ) +
  geom_text_repel(
    data = filter(ssa_test_a, ssaaer == 0.99, name == "QUAMICHAN LAKE; CENTRE"),
    colour = "red",
    label = "Recommended default = 0.99",
    nudge_y = .001
  ) +
  labs(
    title = "Ratio of phototoxic:narcotic benchmark for Anthracene at 3 sites in B.C. for a range of\nAeorosol Single Scattering Albedo",
    x = "Aeorosol Single Scattering Albedo",
    y = "Ratio of phototoxic:narcotic benchmark"
  )
```

### Effect of varying Aerosol Single Scattering Albedo on phototoxic benchmark and ratio of phototoxic:narcotic benchmark for Benzo[a]pyrene

```{r hide_src=TRUE}
ssa_test_b <- multi_pb(sites,
                        pah = "Benzo[a]pyrene",
                        varying = "ssaaer",
                        vals = ssa_vals,
                        DOC = 5,
                        o3_tc = 300,
                        tauaer = 0.235)

ggplot(ssa_test_b, aes(x = ssaaer, y = phototoxic_benchmark)) +
  geom_point() +
  facet_wrap(vars(name), ncol = 1) +
  geom_point(
    data = filter(ssa_test_b, ssaaer == 0.99),
    colour = "red"
  ) +
  geom_text_repel(
    data = filter(ssa_test_b, ssaaer == 0.99, name == "QUAMICHAN LAKE; CENTRE"),
    colour = "red",
    label = "Recommended default = 0.99",
    nudge_y = .001
  ) +
  labs(
    title = "Phototoxic benchmark for Benzo[a]pyrene at 3 sites in B.C. for a range of\nAeorosol Single Scattering Albedo",
    x = "Aeorosol Single Scattering Albedo",
    y = "Phototoxic benchmark (Î¼g/L)"
  )

ggplot(ssa_test_b, aes(x = ssaaer, y = p_n_ratio)) +
  geom_point() +
  facet_wrap(vars(name), ncol = 1) +
  geom_point(
    data = filter(ssa_test_b, ssaaer == 0.99),
    colour = "red"
  ) +
  geom_text_repel(
    data = filter(ssa_test_b, ssaaer == 0.99, name == "QUAMICHAN LAKE; CENTRE"),
    colour = "red",
    label = "Recommended default = 0.99",
    nudge_y = .0005
  ) +
  labs(
    title = "Ratio of phototoxic:narcotic benchmark for Benzo[a]pyrene at 3 sites in B.C. for a range of\nAeorosol Single Scattering Albedo",
    x = "Aeorosol Single Scattering Albedo",
    y = "Ratio of phototoxic:narcotic benchmark"
  )
```

## Latitude (`lat`)

Latitude has a strong effect on the angle of the sun, which will in turn affect
the light penetration through water, thus it is important to investigate the
sensitivity of the calculation of phototoxic benchmark to variation in latitude.

For this analysis, we will maintain the longitudes associated with each site,
and make "mock" sites by creating a range of latitudes from 48 to 72 degrees N
and pair those with the longitudes of the original sites.

```{r}
lat_vals <- seq(48, 72, by = 2)
```

### Effect of varying Latitude on phototoxic benchmark and ratio of phototoxic:narcotic benchmark for Anthracene

```{r hide_src=TRUE}
lat_test_a <- multi_pb(sites,
                        pah = "Anthracene",
                        varying = "lat",
                        vals = lat_vals,
                        DOC = 5,
                        o3_tc = 300,
                        tauaer = 0.235)

ggplot(lat_test_a, aes(x = lat, y = phototoxic_benchmark)) +
  geom_point() +
  facet_wrap(vars(lon), ncol = 1, labeller = as_labeller(\(x) paste("Longitude: ", x))) +
  labs(
    title = "Phototoxic benchmark for Anthracene at 3 sites in B.C. for a range of latitudes",
    x = "Latitude",
    y = "Phototoxic benchmark (Î¼g/L)"
  )

ggplot(lat_test_a, aes(x = lat, y = p_n_ratio)) +
  geom_point() +
  facet_wrap(vars(lon), ncol = 1, labeller = as_labeller(\(x) paste("Longitude: ", x))) +
  labs(
    title = "Ratio of phototoxic:narcotic benchmark for Anthracene at 3 sites in B.C. for a range of latitudes",
    x = "Latitude",
    y = "Ratio of phototoxic:narcotic benchmark"
  )
```

### Effect of varying Latitude on phototoxic benchmark and ratio of phototoxic:narcotic benchmark for Benzo[a]pyrene

```{r hide_src=TRUE}
lat_test_b <- multi_pb(sites,
                        pah = "Benzo[a]pyrene",
                        varying = "lat",
                        vals = lat_vals,
                        DOC = 5,
                        o3_tc = 300,
                        tauaer = 0.235)

ggplot(lat_test_b, aes(x = lat, y = phototoxic_benchmark)) +
  geom_point() +
  facet_wrap(vars(lon), ncol = 1, labeller = as_labeller(\(x) paste("Longitude: ", x))) +
  labs(
    title = "Phototoxic benchmark for Benzo[a]pyrene at 3 sites in B.C. for a range of latitudes",
    x = "Latitude",
    y = "Phototoxic benchmark (Î¼g/L)"
  )

ggplot(lat_test_b, aes(x = lat, y = p_n_ratio)) +
  geom_point() +
  facet_wrap(vars(lon), ncol = 1, labeller = as_labeller(\(x) paste("Longitude: ", x))) +
  labs(
    title = "Ratio of phototoxic:narcotic benchmark for Benzo[a]pyrene at 3 sites in B.C. for a range of\nlatitudes",
    x = "Latitude",
    y = "Ratio of phototoxic:narcotic benchmark"
  )
```

## Elevation (`elev_m`)

We can test the effect of elevation on phototoxic benchmark by varying elevation from sea level
to 1000m above sea level. We will keep the latitude and longitude associated
with each site.

```{r}
elevation_vals <- seq(0, 1000, by = 100)
```

### Effect of varying Elevation on phototoxic benchmark and ratio of phototoxic:narcotic benchmark for Anthracene

```{r hide_src=TRUE}
elev_test_a <- multi_pb(sites,
                         pah = "Anthracene",
                         varying = "elev_m",
                         vals = elevation_vals,
                         DOC = 5,
                         o3_tc = 300,
                         tauaer = 0.235)

ggplot(elev_test_a, aes(x = elev_m, y = phototoxic_benchmark)) +
  geom_point() +
  facet_wrap(vars(name), ncol = 1) +
  labs(
    title = "Phototoxic benchmark for Anthracene at 3 sites in B.C. for a range of elevations",
    x = "Elevation (m)",
    y = "Phototoxic benchmark (Î¼g/L)"
  )

ggplot(elev_test_a, aes(x = elev_m, y = p_n_ratio)) +
  geom_point() +
  facet_wrap(vars(name), ncol = 1) +
  labs(
    title = "Ratio of phototoxic:narcotic benchmark for Anthracene at 3 sites in B.C. for a range of\nelevations",
    x = "Elevation (m)",
    y = "Ratio of phototoxic:narcotic benchmark"
  )
```

### Effect of varying Elevation on phototoxic benchmark and ratio of phototoxic:narcotic benchmark for Benzo[a]pyrene

```{r hide_src=TRUE}
elev_test_b <- multi_pb(sites,
                         pah = "Benzo[a]pyrene",
                         varying = "elev_m",
                         vals = elevation_vals,
                         DOC = 5,
                         o3_tc = 300,
                         tauaer = 0.235)

ggplot(elev_test_b, aes(x = elev_m, y = phototoxic_benchmark)) +
  geom_point() +
  facet_wrap(vars(name), ncol = 1) +
  labs(
    title = "Phototoxic benchmark for Benzo[a]pyrene at 3 sites in B.C. for a range of elevations",
    x = "Elevation (m)",
    y = "Phototoxic benchmark (Î¼g/L)"
  )

ggplot(elev_test_b, aes(x = elev_m, y = p_n_ratio)) +
  geom_point() +
  facet_wrap(vars(name), ncol = 1) +
  labs(
    title = "Ratio of phototoxic:narcotic benchmark for Benzo[a]pyrene at 3 sites in B.C. for a range of\nelevations",
    x = "Elevation (m)",
    y = "Ratio of phototoxic:narcotic benchmark"
  )
```

## Water depth (`depth_m`)

The depth at which the photoxicity of a PAH is determined
should be a conservative one - i.e., a depth at which sensitive species/life stages exist and thus would be exposed. The default is set at 0.25m.

We can test the effect of depth on phototoxic benchmark by varying
elevation from water level to 1m below the surface:

```{r}
depth_vals <- c(seq(0, 1, by = 0.1), 0.25)
```

### Effect of varying Water Depth on phototoxic benchmark and ratio of phototoxic:narcotic benchmark for Anthracene

```{r hide_src=TRUE}
depth_test_a <- multi_pb(sites,
                          pah = "Anthracene",
                          varying = "depth_m",
                          vals = depth_vals,
                          DOC = 5,
                          o3_tc = 300,
                          tauaer = 0.235)

ggplot(depth_test_a, aes(x = depth_m, y = phototoxic_benchmark)) +
  geom_point() +
  geom_point(
    data = filter(depth_test_a, depth_m == 0.25),
    colour = "red"
  ) +
  geom_text_repel(
    data = filter(depth_test_a, depth_m == 0.25, name == "QUAMICHAN LAKE; CENTRE"),
    colour = "red",
    label = "Recommended default = 0.25m",
    nudge_x = -0.05,
    nudge_y = 5
  ) +
  facet_wrap(vars(name), ncol = 1) +
  labs(
    title = "Phototoxic benchmark for Anthracene at 3 sites in B.C. for a range of water depths",
    x = "Water Depth (m)",
    y = "Phototoxic benchmark (Î¼g/L)"
  )

ggplot(depth_test_a, aes(x = depth_m, y = p_n_ratio)) +
  geom_point() +
  geom_point(
    data = filter(depth_test_a, depth_m == 0.25),
    colour = "red"
  ) +
  geom_text_repel(
    data = filter(depth_test_a, depth_m == 0.25, name == "QUAMICHAN LAKE; CENTRE"),
    colour = "red",
    label = "Recommended default = 0.25m",
    nudge_x = -0.05,
    nudge_y = 0.1
  ) +
  facet_wrap(vars(name), ncol = 1) +
  labs(
    title = "Ratio of phototoxic:narcotic benchmark for Anthracene at 3 sites in B.C. for a range of\nwater depths",
    x = "Water Depth (m)",
    y = "Ratio of phototoxic:narcotic benchmark"
  )
```

### Effect of varying Water Depth on phototoxic benchmark and ratio of phototoxic:narcotic benchmark for Benzo[a]pyrene

```{r hide_src=TRUE}
depth_test_b <- multi_pb(sites,
                          pah = "Benzo[a]pyrene",
                          varying = "depth_m",
                          vals = depth_vals,
                          DOC = 5,
                          o3_tc = 300,
                          tauaer = 0.235)

ggplot(depth_test_b, aes(x = depth_m, y = phototoxic_benchmark)) +
  geom_point() +
  geom_point(
    data = filter(depth_test_b, depth_m == 0.25),
    colour = "red"
  ) +
  geom_text_repel(
    data = filter(depth_test_b, depth_m == 0.25, name == "QUAMICHAN LAKE; CENTRE"),
    colour = "red",
    label = "Recommended default = 0.25m",
    nudge_x = -0.05,
    nudge_y = 0.1
  ) +
  facet_wrap(vars(name), ncol = 1) +
  labs(
    title = "Phototoxic benchmark for Benzo[a]pyrene at 3 sites in B.C. for a range of water depths",
    x = "Water Depth (m)",
    y = "Phototoxic benchmark (Î¼g/L)"
  )

ggplot(depth_test_b, aes(x = depth_m, y = p_n_ratio)) +
  geom_point() +
  geom_point(
    data = filter(depth_test_b, depth_m == 0.25),
    colour = "red"
  ) +
  geom_text_repel(
    data = filter(depth_test_b, depth_m == 0.25, name == "QUAMICHAN LAKE; CENTRE"),
    colour = "red",
    label = "Recommended default = 0.25m",
    nudge_x = -0.05,
    nudge_y = 0.05
  ) +
  facet_wrap(vars(name), ncol = 1) +
  labs(
    title = "Ratio of phototoxic:narcotic benchmark for Benzo[a]pyrene at 3 sites in B.C. for a range of\nwater depths",
    x = "Water Depth (m)",
    y = "Ratio of phototoxic:narcotic benchmark"
  )
```


## Radiative transfer scheme (`nstr`)

Number of streams for radiative transfer calculations:

* If `nstr` < 2, uses 2-stream delta-Eddington (faster)
* if `nstr` >= 2, uses n-stream discrete ordinates (more accurate: must be even number, maximum = 32)

```{r, hide_src=TRUE, cache=TRUE, fig.height=6}
out <- multi_pb(sites,
          pah = "Anthracene",
          varying = "nstr",
          vals = c(-2, 4),
          DOC = 5,
          o3_tc = 300,
          tauaer = 0.235) |>
  mutate(nstr = factor(nstr))

ggplot(out, aes(x = name, y = phototoxic_benchmark, colour = nstr)) +
  geom_point() +
  labs(
    title = "Calculated phototoxic benchmark using two different TUV methods for radiative transfer",
    x = "Site"
  ) +
  theme(axis.text.x = element_text(angle = 355))

ggplot(out, aes(x = name, y = timing, colour = nstr)) +
  geom_point() +
  labs(
    title = "Execution time of TUV model runs with different methods for radiative transfer",
    x = "Site",
    y = "Execution Time (s)"
  ) +
  theme(axis.text.x = element_text(angle = 355))
```

We can see that there are small differences in the calculated phototoxic benchmark when using different values of `nstr`, with the values when calculated using `nstr = -2` being on average
`r round(out |> group_by(name) |> summarize(diff = (phototoxic_benchmark[nstr == 4] - phototoxic_benchmark[nstr == -2]) / phototoxic_benchmark[nstr == 4]) |> pull(diff) |> mean() * 100, 2)`% lower than when using `nstr = 4`. Using `nstr = 4` however, is about `r round(out |> group_by(name) |> summarize(diff = (timing[nstr == 4] / timing[nstr == -2])) |> pull(diff) |> mean())` times slower than using `nstr = -2`. In most cases, unless high precision is required, it is likely that using the much faster `nstr = -2` should be the default.

## Methylated Polycyclic Aromatic Hydrocarbons

Absorption spectra for methylated PAHs are not all available from published sources, so it would be beneficial to be able to approximate them for calculating
water quality guidelines. In this section we will look at the feasibilty of using the absorbance spectra from a methylated PAH's parent compound as a proxy for the specific spectra. We will do this by calculating the phototoxic benchmark for various methylated PAHs for which we have absorbance spectra, and comparing these values to those calculated using the absorbance spectra of the parent PAH.

For these comparisons we will limit it to just one site in the Okanagan.

```{r}
okanagan <- filter(sites, emsid == "0500236")
```

```{r, hide_src=TRUE}
# Define some functions:
ma_plot <- function(chemicals) {
  pahwq:::molar_absorption |>
    filter(chemical %in% tolower(chemicals)) |>
    mutate(
      chemical = factor(chemical, levels = tolower(chemicals))
    ) |>
    ggplot(aes(x = wavelength, y = molar_absorption, colour = chemical)) +
    geom_point() +
    labs(
      title = paste0("Molar absorption of ", chemicals[1], " and methylated derivatives"),
      y = "Molar absorption (L/mol/cm)"
    )
}

nb_plot <- function(chemicals) {
  pahwq:::nlc50_lookup |>
    filter(chemical %in% tolower(chemicals)) |>
    mutate(
      narcotic_benchmark = vapply(chemical, narcotic_benchmark, FUN.VALUE = 1),
      chemical = factor(chemical, levels = tolower(chemicals))
    ) |>
    ggplot(aes(x = log_kow, y = narcotic_benchmark, colour = chemical)) +
    geom_point() +
    labs(
      title = paste0("Narcotic benchmark of ", chemicals[1], " and its methylated derivatives"),
      y = "Narcotic benchmark (ug/L)"
    )
}

pb_surrogates <- function(chemicals) {
  multi_pb(
  okanagan,
  site = "name",
  pah = chemicals,
  varying = "Kd_ref",
  vals = c(1, 150),
  depth_m = 0.25
) |>
  group_by(Kd_ref) |>
  rowwise() |>
  mutate(
    Pabs_parent = p_abs(tuv_res, tolower(chemicals[1])),
    pb_parent = phototoxic_benchmark(Pabs_parent, PAH)
  ) |>
  ungroup() |>
  mutate(
    PAH = factor(PAH, levels = chemicals)
  ) |>
  pivot_longer(cols = c(phototoxic_benchmark, pb_parent),
               names_to = "abs_spectra",
               values_to = "phototoxic_benchmark") |>
  mutate(
    p_n_ratio = phototoxic_benchmark / narcotic_benchmark,
    abs_spectra = case_when(
      abs_spectra == "phototoxic_benchmark" ~ "specific",
      abs_spectra == "pb_parent" ~ "parent"
    )
  )
}

pb_plot <- function(df) {

ggplot(df, aes(x = PAH, y = phototoxic_benchmark, colour = abs_spectra)) +
  geom_point() +
  facet_wrap(vars(Kd_ref), ncol = 1, scales = "free_y",
             labeller = as_labeller(function(x) paste0("Kd(305) = ", x))) +
  geom_text_repel(
    data = filter(df, PAH != chemicals[1]),
    mapping = aes(label = round(p_n_ratio, 3)),
    show.legend = FALSE
  ) +
  labs(
    title = paste0("Comparison of phototoxic benchmark of methylated ", chemicals[1], ", using\nspecific absorption spectra vs absorption spectra of parent compound"),
    y = "Phototoxic benchmark (ug/L)",
    colour = "Absorption spectra used",
    caption = "*Text labels indicate the ratio of phototoxic:narcotic benchmark"
  )
}

```


### Napthalene

To start, we can examine the molar absorption specra of three methylated Naphthalene derivatives, as compared to unmethylated Naphthalene:

```{r, hide_src=TRUE, fig.height=6}
chemicals <- c("Naphthalene", "1-Methylnaphthalene", "2,6-Dimethylnaphthalene", "1,6,7-Trimethylnaphthalene")

ma_plot(chemicals)
```

We can see in general that more highly methylated compounds have greater absorption, though it is not consistent across the spectrum.

We can also examine the properties and narcotic toxicity of Naphthalene and its methylated variants:

```{r, hide_src=TRUE, fig.height=6}
nb_plot(chemicals)
```

We can see here that 1,6,7-Trimethylnaphthalene has the highest log(KoW) and lowest narcotic benchmark.

Finally, we can compare the phototoxicity (phototoxic benchmark) of Naphthalene and its methylated
derivatives at the Okanagan site, at the two values of $k_{d,305}$ (1 and 150).

```{r, hide_src=TRUE}
df <- pb_surrogates(chemicals)
pb_plot(df) +
  theme(axis.text.x = element_text(angle = 355))
```

Here we can see that a lower phototoxic benchmark (higher phototoxicity) is calculated
for all methylated naphthalenes when the absorption spectra of the methylated compound is used, as compared to when that of the parent compound (Naphthalene) is used. This difference is much greater when there is higher light penetration (lower Kd) - which is to be expected, as when there is less light penetration the phototoxic benchmark
approaches the narcotic benchmark so the absorption spectra of the chemical becomes irrelevant.

### Phenanthrene

As above, we can visualize the absorption spectra of Phenanthrene and 9-Methylphenanthrene:

```{r, hide_src=TRUE, fig.height=6}
chemicals <- c("Phenanthrene", "9-Methylphenanthrene")

ma_plot(chemicals)
```

Here we see that the parent compound Phenanthrene has much higher absorption than
9-Methylphenanthrene at the low end of the light spectrum.

```{r, hide_src=TRUE, fig.height=6}
nb_plot(chemicals)
```

9-Methylphenanthrene has higher narcotic toxicity (lower narcotic benchmark) than its parent
Phenanthrene.

```{r, hide_src=TRUE}
phototoxic_benchark_df_phenanthrene <- pb_surrogates(chemicals)
pb_plot(phototoxic_benchark_df_phenanthrene)
```
Using the parent molar absorption spectra for 9-Methylphenanthrene gives a more conservative phototoxic benchmark than using the specific spectra for the methylated compound. This is different than the case for Naphthalene.

Similar to Naphthalene, this only appears to be substantially different at high light penetration (low Kd(305)).

### Anthracene

As above, we can visualize the absorption spectra of Anthracene and 9-Methylanthracene:

```{r, hide_src=TRUE, fig.height=6}
chemicals <- c("Anthracene", "9-Methylanthracene")
ma_plot(chemicals)
```

The absorption spectra for both Anthracene and 9-Methylanthracene is very
broad compared to that of the other PAH compounds.

```{r, hide_src=TRUE, fig.height=6}
nb_plot(chemicals)
```

9-Methylanthracene has higher narcotic toxicity (lower narcotic benchmark) than its parent
Anthracene.

```{r, hide_src=TRUE}
phototoxic_benchmark_df_anthracene <- pb_surrogates(chemicals)
pb_plot(phototoxic_benchmark_df_anthracene) +
  theme(axis.text.x = element_text(angle = 355))
```
Using the parent absorption spectra for 9-Methylanthracene results in a less
conservative phototoxic benchmark than if the specific absorption spectra is used. Contrary to that seen in the other PAHs above, this effect is still quite pronounced at high
light attenuation (Kd(305) = 150), presumably because of the broad absorption spectra exhibited by both Anthracene and 9-Methylanthracene.

### Pyrene

```{r, hide_src=TRUE, fig.height=6}
chemicals <- c("Pyrene", "1-Methylpyrene")

ma_plot(chemicals)
```


```{r, hide_src=TRUE, fig.height=6}
nb_plot(chemicals)
```

9-Methylpyrene has higher narcotic toxicity (lower narcotic benchmark) than its parent
Pyrene.

```{r, hide_src=TRUE}
phototoxic_benchmark_df_pyrene <- pb_surrogates(chemicals)
pb_plot(phototoxic_benchmark_df_pyrene)
```
Using the parent absorption spectra for 9-Methylpyrene results in a less
conservative phototoxic benchmark than if the specific absorption spectra is used.

### Chrysene

```{r, hide_src=TRUE, fig.height=6}
chemicals <- c("Chrysene", "1-Methylchrysene")
ma_plot(chemicals)
```

```{r, hide_src=TRUE, fig.height=6}
nb_plot(chemicals)
```

Similar to the other methylated PAHs, 1-Methylchrysene has a lower narcotic benchmark (higher narcotic toxicity) than its parent Chrysene.

```{r, hide_src=TRUE}
phototoxic_benchmark_df_chrysene <- pb_surrogates(chemicals)
pb_plot(phototoxic_benchmark_df_chrysene)
```
The results for Chrysene and 1-Methylchrysene show that the effect of using the absorbance spectra of a parent compound can be different depending on the light
intensity/attenuation. At low light attenuation (Kd(305) = 1), the specific spectra gives a more conservative phototoxic benchmark for 1-Methylchrysene than that obtained using the
spectra from Chrysene, while the opposite is true at high light attenuation (Kd(305) = 100).

These results show that substituting the absorption spectra of a parent compound when that of a specific methylated PAH is unknown is possible, but also somewhat
complicated. Differences in absorption spectra between methylated PAHs and that of their parent was quite variable across compounds. The difference in the resulting calculated phototoxic benchmark is generally quite small and thus it is likely reasonable to use the absorption spectra of a parent when it is not available for a methylated PAH. However, the magnitude and even the sign of the difference is highly variable. In all cases examined, the narcotic benchmark of the methylated compounds was lower than that of the parent compounds (i.e., they are more toxic), however the final phototoxicity (phototoxic benchmark) depended on the differences in the absorption spectra and, in some cases, the level of light attenuation.

## References
