---
title: "Sensitivity Analysis"
bibliography: references.bib
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 8,
  fig.height = 8
)

local({
  hook_old <- knitr::knit_hooks$get("source")  # save the old hook
  knitr::knit_hooks$set(source = function(x, options) {
    if (!is.null(options$hide_src) && options$hide_src) {
      return(
        c("<details>
          <summary style='font-size:0.8em; margin-bottom:0; color: #666;'>
          Show/Hide Code
          </summary>\n", 
          hook_old(x, options), 
          "</details>")
      )
    }
    hook_old(x, options)
  })
})
```

```{r}
#| include: false
library(sf)
library(mapview)
library(rems)
library(dplyr)
library(gt)
library(ggplot2)
library(ggrepel)
library(latex2exp)

theme_set(theme_bw())

sites_sf <- structure(list(emsid = c("0400390", "0500236", "E207466"), name = c("CHARLIE L DEEP STATION 1.2 KM EAST OF PARK", 
"OKANAGAN L D/S KELOWNA STP (DEEP)", "QUAMICHAN LAKE; CENTRE"
), lon = c(-120.9642, -119.5134, -123.6625), lat = c(56.3125, 
49.8614, 48.8003), elev_m = c(693, 342, 25), date = structure(c(1629823800, 
1631127600, 1629918000), tzone = "Etc/GMT+8", class = c("POSIXct", 
"POSIXt")), DOC = c(14, 4.26, 6.61), doc_min = c(0.96, 4.06, 
6.37), doc_max = c(15.4, 5.17, 11.8), geometry = structure(list(
    structure(c(-120.9642, 56.3125), class = c("XY", "POINT", 
    "sfg")), structure(c(-119.5134, 49.8614), class = c("XY", 
    "POINT", "sfg")), structure(c(-123.6625, 48.8003), class = c("XY", 
    "POINT", "sfg"))), class = c("sfc_POINT", "sfc"), precision = 0, crs = structure(list(
    input = "EPSG:4326", wkt = "GEOGCRS[\"WGS 84\",\n    ENSEMBLE[\"World Geodetic System 1984 ensemble\",\n        MEMBER[\"World Geodetic System 1984 (Transit)\"],\n        MEMBER[\"World Geodetic System 1984 (G730)\"],\n        MEMBER[\"World Geodetic System 1984 (G873)\"],\n        MEMBER[\"World Geodetic System 1984 (G1150)\"],\n        MEMBER[\"World Geodetic System 1984 (G1674)\"],\n        MEMBER[\"World Geodetic System 1984 (G1762)\"],\n        MEMBER[\"World Geodetic System 1984 (G2139)\"],\n        ELLIPSOID[\"WGS 84\",6378137,298.257223563,\n            LENGTHUNIT[\"metre\",1]],\n        ENSEMBLEACCURACY[2.0]],\n    PRIMEM[\"Greenwich\",0,\n        ANGLEUNIT[\"degree\",0.0174532925199433]],\n    CS[ellipsoidal,2],\n        AXIS[\"geodetic latitude (Lat)\",north,\n            ORDER[1],\n            ANGLEUNIT[\"degree\",0.0174532925199433]],\n        AXIS[\"geodetic longitude (Lon)\",east,\n            ORDER[2],\n            ANGLEUNIT[\"degree\",0.0174532925199433]],\n    USAGE[\n        SCOPE[\"Horizontal component of 3D system.\"],\n        AREA[\"World.\"],\n        BBOX[-90,-180,90,180]],\n    ID[\"EPSG\",4326]]"), class = "crs"), n_empty = 0L, bbox = structure(c(xmin = -123.6625, 
ymin = 48.8003, xmax = -119.5134, ymax = 56.3125), class = "bbox"))), row.names = c(NA, 
-3L), sf_column = "geometry", agr = structure(c(emsid = NA_integer_, 
name = NA_integer_, lon = NA_integer_, lat = NA_integer_, elev_m = NA_integer_, 
date = NA_integer_, DOC = NA_integer_, doc_min = NA_integer_, 
doc_max = NA_integer_), levels = c("constant", "aggregate", "identity"
), class = "factor"), class = c("sf", "tbl_df", "tbl", "data.frame"
))

`%||%` <- function(x, y) if (is.null(x)) y else x
```

This report is to document the sensitivity of the TUV model, and the subsequent
calculation of water quality guidelines for PAHs, to variation in certain TUV
parameters.

## Input parameters

The input parameters for the calculation of light attenuation through water for
the TUV model are: 

  * Vertical Attenuation Coefficient at wavelength `wvl`: $kvat = k_de^{-Sk(ref\_wvl-wvl)}$
      - **`Kd`** Light attenuation coefficient (set directly or calculated from DOC)
      - **`Sk`** (default `0.018`)
      - **`ref_wvl`** Reference wavelength for `Kd` (default `305`)
  * **`depth_m`** Water depth in m
  * **`lat`** Latitude in decimal degrees
  * **`lon`** Longitude in decimal degrees
  * **`elev_km`** Surface elevation, km above sea level
  * **`year`** Year
  * **`month`** Month
  * **`day`** Day
  * **`tzone`** Timezone  Local Time - UTC (default `0`; UTC)
  * **`tstart`** Start time, hours local time (default `0`)
  * **`tstop`** Stop time, hours local time (default `23`)
  * **`tsteps`** Number of time steps (default `24`)
  * **`albedo`** Surface albedo (default `0.07`)
  * **`o3_tc`** Ozone column, Dobson Units (DU) - Looked up in climatology (recommended default `300`)
  * **`so2_tc`** SO2 column, DU (default `0`)
  * **`no2_tc`** NO2 column, DU (default `0`)
  * **`taucld`** Cloud optical depth (default `0`)
  * **`zbase`** Cloud base, km (default `4`)
  * **`ztop`** Cloud top, km (default `5`)
  * **`tauaer`** Aerosol optical depth at 550 nm - Looked up in climatology (recommended default `0.235`)
  * **`ssaaer`** Aerosol single scattering albedo (default `0.990`)
  * **`alpha`** Aerosol Angstrom exponent(default `1.0`) 
  * **`wvl_start`** Starting wavelength, nm (default `279.5`)
  * **`wvl_end`** End wavelength, nm (default `420.5`)
  * **`wvl_steps`** Number of wavelength intervals (default `141`)
  * **`nstr`** TUV run type; use -2 for fast, 4 for slightly more accurate (default `-2`)

Those currently under consideration for sensitivity analysis are:

  * **`Kd`** Light attenuation coefficient - includes use of **`DOC`** for calculating `Kd`
  * **`albedo`** Surface albedo (default `0.07`)
  * **`o3_tc`** Ozone column, Dobson Units (DU) - Looked up in climatology (recommended default `300`)
  * **`tauaer`** Aerosol optical depth at 550 nm - Looked up in climatology (recommended default `0.235`)
  * **`ssaaer`** Aerosol single scattering albedo (default `0.990`)
  * **`lat`** Latitude in decimal degrees
  * **`elev_km`** Surface elevation, km above sea level
  * **`depth_m`** Water depth in m
  * **`nstr`** TUV run type; use -2 for fast, 4 for slightly more accurate (default `-2`)
  
The analyses are conducted at three different sample lakes (Southern Interior, 
Vancouver Island, Northeast), using two PAHs (Anthracene and Benzo[a]pyrene).

For each analysis, we will test a range of reasonable values for the parameter of 
interest, and plot the PLC50 value calculated across that range. To look at the
relative effect of that parameter on the photoxicity of the PAH, we also plot 
the ratio of PLC50:NLC50.

Initially, sensitivity analyses will be univariate (varying the input of interest
while holding the others constant). If there are significant interactions expected
between certain variables, these can be explored.

## Basic usage of the `pahwq` package

To start, we will demonstrate a typical straightforward use case at a single site.

To calculate the acute phototoxic water quality guideline (PLC50) for Anthracene 
at 0.25 m depth in Okanagan Lake on June 21, 2023, with a measured DOC of 5 g/m^3, 
you would use the following code:

First, set up the options for the model run:
```{r}
library(pahwq)

set_tuv_aq_params(
  depth_m = 0.25,
  lat = 49.601632,
  lon = -119.605862,
  elev_km = 0.342,
  DOC = 5,
  date = "2023-06-21",
  tzone = -8,
  albedo = 0.05
)
```

After setting them, you can view the options that will be used by TUV. Some are set via the function
inputs, some are looked up (e.g., `o3_tc`, `tauaer`) or calculated (e.g., `kd(305)` is calculated from the input `DOC`).

```{r}
view_tuv_aq_params()
```

Run the TUV model and get the results. The results show the underwater irradiance 
at the specified depth, at each wavelength and time step.

```{r}
run_tuv()

# Get the results
tuv_res <- get_tuv_results(file = "out_irrad_y")
head(tuv_res)
```

We can inspect and verify the inputs that were used in the model run:

```{r}
tuv_run_params(tuv_res)
```

Next, calculate the site-specific light absorption ($P_{abs}$) for Anthracene from the 
TUV results. The `p_abs()` function uses a lookup table to get the molar 
absorption coefficient table for the specified PAH.

```{r}
(Pabs <- p_abs(tuv_res, "Anthracene"))
```

Finally, calculate PLC50 in ðg/L, supplying the $P_{abs}$ value.

```{r}
plc50(Pabs, pah = "Anthracene")
```

We can compare the PLC50 to the NLC50 to see the effect of the photoxicity of the PAH:

```{r}
nlc50("Anthracene")
```

### A shortcut:

If you don't need to inspect every step of the way, the above process can be completed in two function calls:

```{r}
tuv_res <- tuv(
  depth_m = 0.25,
  lat = 49.601632,
  lon = -119.605862,
  elev_km = 0.342,
  DOC = 5,
  date = "2023-06-21",
  tzone = -8,
  albedo = 0.05
)

plc50(tuv_res, "Anthracene")
```

## Sensitivity Analysis: Setup

We'll start by loading the packages we need and getting some data from B.C. EMS:

```{r setup, hide_src = TRUE}
#| eval: false
library(sf)
library(mapview)
library(rems)
library(dplyr)
library(ggplot2)
library(ggrepel)
library(latex2exp)

con <- connect_historic_db()
db <- attach_historic_data(con)
sites_sf <- db |> 
  filter(
    EMS_ID %in% c("0500236", "E207466", "0400390"),
    # grepl("(CHARLIE L)|(OKANAGAN)|(QUAMICHAN)", MONITORING_LOCATION), 
    COLLECTION_START > as.Date("2020-01-01"), PARAMETER_CODE == "1103"
  ) |> 
  collect() |> 
  st_as_sf(coords = c("LONGITUDE", "LATITUDE"), crs = 4326, remove = FALSE) |> 
  filter(!is.na(RESULT)) |> 
  arrange(desc(COLLECTION_START)) |> 
  group_by(EMS_ID, MONITORING_LOCATION, LONGITUDE, LATITUDE) |> 
  mutate(doc_min = min(RESULT), doc_max = max(RESULT)) |> 
  slice(1) |> 
  select(emsid = EMS_ID, 
         name = MONITORING_LOCATION, 
         lon = LONGITUDE, 
         lat = LATITUDE, 
         date = COLLECTION_START, 
         DOC = RESULT,
         doc_min,
         doc_max) |> 
  left_join(
    tribble(
      ~ emsid, ~elev_m,
      "E207466", 25,
      "0400390", 693,
      "0500236", 342
    ), by = join_by("emsid")
  ) |> 
  ungroup() |> 
  relocate(elev_m, .after = lat)

disconnect_historic_db(con)
```


## Locations
```{r hide_src=TRUE}
mapview(sites_sf, zcol = "name", legend = FALSE)
```

### Okanagan Lake, Okanagan

* EMS ID: 0500236
* Region: Okanagan 
* Latitude: 49.8614 N
* Longitude: 119.5134 W
* Site Depth: 70.1 m 
* Maximum Lake Depth: 232 m
* Lake Elevation: 342 m 
* Lake Surface Area: 350.08 sq km

### Charlie Lake, Peace

* EMS ID: 0400390
* Region: Peace 
* Latitude: 56.3125 N
* Longitude: 120.9642 W
* Site Depth: 13 m 
* Maximum Lake Depth: 13 m
* Lake Elevation: 693 m 
* Lake Surface Area: 17.56 sq km

### Quamichan Lake, Vancouver Island

* EMS ID: E207466
* Region: Vancouver Island 
* Latitude: 48.8003 N
* Longitude: 123.6625 W
* Site Depth: 7 m 
* Maximum Lake Depth: 8 m
* Lake Elevation: 25 m 
* Lake Surface Area: 2.88 sq km

### Prepare the data for comparisons, including setting the date to be the same
for all sites:

```{r hide_src=TRUE}
sites <- sites_sf |> 
  st_drop_geometry() |> 
  mutate(date = "2023-08-01")
gt(sites)
```

## Basic analysis using defaults

To perform the sensitivity analysis, we need to define a function, `multi_plc50()` that allows us to do repeated runs of the TUV model and PLC50 calculation using a range of inputs at multiple sites:

```{r multi-tuv, hide_src=TRUE}
multi_plc50 <- function(df, site = "name", pah, varying, vals = NULL, ...) {
  if (!varying %in% union(names(tuv_aq_defaults()), names(formals(set_tuv_aq_params)))) {
    stop(varying, " is not a valid argument for `set_tuv_aq_params()`")
  }
  
  if (!is.null(vals)) {
    var_df <- data.frame(vals)
    names(var_df) <- varying
    
    df <- df |> 
      select(!any_of(varying)) |> 
      dplyr::cross_join(var_df)
  }
  
  df |> 
    rowwise() |> 
    mutate(
      tuv_res = list(calc_tuv(
          date = date, 
          lat = lat, 
          lon = lon,
          elev_m = elev_m,
          varying = .data[[varying]],
          vary_var = varying,
          ...
        ))
    ) |> 
    cross_join(data.frame(PAH = pah)) |> 
    mutate(
      timing = attr(tuv_res, "timing"),
      Pabs = p_abs(tuv_res, PAH),
      plc50 = plc50(Pabs, pah = PAH),
      plc_nlc_ratio = plc50 / nlc50(PAH)
    ) |> 
    ungroup()
}

calc_tuv <- function(date, lat, lon, elev_m, varying, vary_var, ...) {
  args <- c(
    varying, 
    list(
      depth_m = 0.25,
      date = as.Date(date),
      lat = lat,
      lon = lon,
      elev_km = elev_m / 1000
    ),
    ...
  )
  names(args)[1] <- vary_var
  
  # allow overriding of one of the core args by one supplied in 'varying', 
  # this will keep the first of duplicated argument names, which will
  # be the one in 'varying'
  args <- args[unique(names(args))]
  
  do.call("set_tuv_aq_params", args)
  
  t <- system.time(run_tuv(quiet = TRUE))
  res <- get_tuv_results(file = "out_irrad_y")
  attr(res, "timing") <- unname(t["elapsed"])
  res
}
```

We then use the `multi_plc50()` function to calculate $P_{abs}$, PLC50, and ratio 
of PLC50:NLC50 for Anthracene using recorded DOC values at the three sites. 
This also uses the utility of pahwq to look up ozone column and aerosol optical
depth from climatologies based on latitude, longitude, and month.

```{r hide_src=TRUE}
diff_DOC <- multi_plc50(sites, pah = c("Anthracene", "Naphthalene"), varying = "DOC")
```

```{r}
#| echo: false
diff_DOC |>
  select(-doc_min, -doc_max, -tuv_res) |>
  gt() |>
  data_color(columns = plc50, palette = "Greens")
```

To compare the sites using the same input parameters other than location (lat, lon, elevation), we modify the data to set a constant [DOC] (DOC = 5), and calculate $P_{abs}$, PLC50, and ratio of PLC50:NLC50 for Anthracene using constant DOC = 5: 

```{r hide_src=TRUE}
same_DOC <- sites |>
  mutate(DOC = 5) |>
  multi_plc50(pah = "Anthracene", varying = "DOC")
```

```{r}
#| echo: false
same_DOC |>
  select(-doc_min, -doc_max, -tuv_res) |>
  gt() |>
  data_color(columns = plc50, palette = "Greens")
```

## Light Attenuation Coefficicent (_k~d~_)

The TUV model calculates the light attenuation coefficient $k_d(Î»)$ for each
wavelength, based on a reference $k_d(Î»_{ref})$, where $\lambda_{ref}$ is the reference wavelength. 

$k_d(Î»_{ref})$ at 305nm ($k_{d,305}$) can be estimated from Dissolved Organic
Carbon concentration [DOC] using the equation:

$k_{d,305} = a_{305}[DOC]^{b_{305}} + 0.13$; 
$a_{305}$ = 2.76 and $b_{305}$ = 1.2 from @morris1995; Equation 4-1a in @aris2023

For the sensitivity analysis for $k_d$, it is more interpretable to supply 
a range of $k_d$ values to the TUV model directly, even though in practice it is 
more likely to use [DOC] to estimate $k_d$.

We will test the sensitivity to $k_d$ using sample values from Table 3 (@aris2023), from 0.08 to 275:

```{r hide_src=TRUE}
kd_vals <- seq(0.08, 275, length.out = 20)
```

### Effect of varying _k~d~_ on PLC50 and ratio of PLC50:NLC50 for Anthracene

```{r hide_src=TRUE}
kd_test_a <- multi_plc50(sites, 
                       pah = "Anthracene", 
                       varying = "Kd_ref", 
                       vals = kd_vals, 
                       Kd_wvl = 305,
                       o3_tc = 300,
                       tauaer = 0.235)

ggplot(kd_test_a, aes(x = Kd_ref, y = plc50)) +
  geom_point() + 
  facet_wrap(vars(name), ncol = 1) +
  labs(
    title = TeX(r"(PLC50 for Anthracene at 3 sites in B.C. for a range of $k_d(305)$)"), 
    x = TeX(r"($k_d(305)$)"),
    y = "PLC50 (Î¼g/L)"
  )

ggplot(kd_test_a, aes(x = Kd_ref, y = plc_nlc_ratio)) +
  geom_point() + 
  facet_wrap(vars(name), ncol = 1) +
  labs(
    title = TeX(r"(Ratio of PLC50:NLC50 for Anthracene at 3 sites in B.C. for a range of $k_d(305)$)"), 
    x = TeX(r"($k_d(305)$)"),
    y = "Ratio of PLC50:NLC50"
  )
```

For Anthracene, the ratio of PLC50:NLC50 approaches 1 when $k_d(305)$ is ~ 
250 --- in other words at that level of $k_d$, PLC50 is nearly the same as NLC50, 
indicating that the light is attenuated to such an extent that the phototoxicity 
of Anthracene is not activated at the light levels when $k_d(305)$ >= 250.

According to the above equation for estimating $k_d$ from [DOC]: 

$250 = 2.76[DOC]^{1.23} + 0.13$

[DOC] ~= 40

Therefore $k_d(305)$ of ~250 is representative of light attenuation when [DOC] 
is ~40. This is a significantly higher DOC concentration than observed in our 
three sample lakes, and is outside the range for which the above equation is 
recommended (it is recommended for [DOC] between 0.2 and 23 mg/L).

### Effect of varying _k~d~_ on PLC50 and ratio of PLC50:NLC50 for Benzo[a]pyrene

We will use the same input values as we did for Anthracene:

```{r hide_src=TRUE}
kd_test_b <- multi_plc50(sites, 
                       pah = "Benzo[a]pyrene", 
                       varying = "Kd_ref", 
                       vals = kd_vals,
                       Kd_wvl = 305,
                       o3_tc = 300,
                       tauaer = 0.235)

ggplot(kd_test_b, aes(x = Kd_ref, y = plc50)) +
  geom_point() + 
  facet_wrap(vars(name), ncol = 1) +
  labs(
    title = TeX(r"(PLC50 for Benzo\[a\]pyrene at 3 sites in B.C. for a range of $k_d(305)$)"), 
    x = TeX(r"($k_d(305)$)"),
    y = "PLC50 (Î¼g/L)"
  )

ggplot(kd_test_b, aes(x = Kd_ref, y = plc_nlc_ratio)) +
  geom_point() + 
  facet_wrap(vars(name), ncol = 1) +
  labs(
    title = TeX(r"(Ratio of PLC50:NLC50 for Benzo\[a\]pyrene at 3 sites in B.C. for a range of $k_d(305)$)"), 
    x = TeX(r"($k_d(305)$)"),
    y = "Ratio of PLC50:NLC50"
  )
```

## Surface Albedo (`albedo`)

We test surface albedo values from 0.05 to 0.1 (typical for water), and include 
the suggested default of 0.7:

```{r}
albedo_vals <- c(seq(0.05, 0.1, length.out = 10), 0.07)
```

For this and the remaining analyses, we will set DOC = 5 for all sites to ensure 
a constant $k_d$ value. We can then test the range of surface albedo values at 
all three sites for Anthracene and Benzo[a]pyrene.

### Effect of varying Surface Albedo on PLC50 and ratio of PLC50:NLC50 for Anthracene

```{r hide_src=TRUE}
albedo_test_a <- multi_plc50(sites, 
                           pah = "Anthracene", 
                           varying = "albedo", 
                           vals = albedo_vals,
                           DOC = 5,
                           o3_tc = 300,
                           tauaer = 0.235)

ggplot(albedo_test_a, aes(x = albedo, y = plc50)) +
  geom_point() + 
  facet_wrap(vars(name), ncol = 1) +
  geom_point(
    data = filter(albedo_test_a, albedo == 0.07),
    colour = "red"
  ) + 
  geom_text_repel(
    data = filter(albedo_test_a, albedo == 0.07, name == "QUAMICHAN LAKE; CENTRE"),
    colour = "red", 
    label = "Recommended default = 0.07",
    nudge_x = 0.001,
    nudge_y = 0.05
  ) +
  labs(
    title = "PLC50 for Anthracene at 3 sites in B.C. for a range of Surface Albedo", 
    x = "Surface Albedo",
    y = "PLC50 (Î¼g/L)"
  )

ggplot(albedo_test_a, aes(x = albedo, y = plc_nlc_ratio)) +
  geom_point() + 
  facet_wrap(vars(name), ncol = 1) +
  geom_point(
    data = filter(albedo_test_a, albedo == 0.07),
    colour = "red"
  ) + 
  geom_text_repel(
    data = filter(albedo_test_a, albedo == 0.07, name == "QUAMICHAN LAKE; CENTRE"),
    colour = "red", 
    label = "Recommended default = 0.07",
    nudge_x = 0.001,
    nudge_y = 0.0005
  ) +
  labs(
    title = "Ratio of PLC50:NLC50 for Anthracene at 3 sites in B.C. for a range of\nSurface Albedo", 
    x = "Surface Albedo",
    y = "Ratio of PLC50:NLC50"
  )
```

### Effect of varying Surface Albedo on PLC50 and ratio of PLC50:NLC50 for Benzo[a]pyrene

```{r hide_src=TRUE}
albedo_test_b <- multi_plc50(sites, 
                           pah = "Benzo[a]pyrene", 
                           varying = "albedo", 
                           vals = albedo_vals,
                           DOC = 5,
                           o3_tc = 300,
                           tauaer = 0.235)

ggplot(albedo_test_b, aes(x = albedo, y = plc50)) +
  geom_point() + 
  facet_wrap(vars(name), ncol = 1) +
  geom_point(
    data = filter(albedo_test_b, albedo == 0.07),
    colour = "red"
  ) + 
  geom_text_repel(
    data = filter(albedo_test_b, albedo == 0.07, name == "QUAMICHAN LAKE; CENTRE"),
    colour = "red", 
    label = "Recommended default = 0.07",
    nudge_x = 0.001,
    nudge_y = 0.0005
  ) +
  labs(
    title = "PLC50 for Benzo[a]pyrene at 3 sites in B.C. for a range of Surface Albedo", 
    x = "Surface Albedo",
    y = "PLC50 (Î¼g/L)"
  )

ggplot(albedo_test_b, aes(x = albedo, y = plc_nlc_ratio)) +
  geom_point() + 
  facet_wrap(vars(name), ncol = 1) +
  geom_point(
    data = filter(albedo_test_b, albedo == 0.07),
    colour = "red"
  ) + 
  geom_text_repel(
    data = filter(albedo_test_b, albedo == 0.07, name == "QUAMICHAN LAKE; CENTRE"),
    colour = "red", 
    label = "Recommended default = 0.07",
    nudge_x = 0.001,
    nudge_y = 0.0003
  ) +
  labs(
    title = "Ratio of PLC50:NLC50 for Benzo[a]pyrene at 3 sites in B.C. for a range of\nSurface Albedo", 
    x = "Surface Albedo",
    y = "Ratio of PLC50:NLC50"
  )
```

## Ozone Column (`o3_tc`)

The pahwq package currently uses average monthly ozone column data from 1980-1991
from @fortuin1998, which is bundled with the TUV model. Based on 
latitude and longitude and month, the ozone column value is looked up and supplied
to the TUV model. This default behaviour can be overridden by supplying a value to 
the `o3_tc` argument in the `set_tuv_aq_params` function. In the absence of
climatological data, the recommended default value is 300 DU (@aris2023).

To test the sensitivity of PLC50 to ozone column, a range of values is tested from
280-420 DU, which are typical values in middle to Northern latitudes. See [NASA's 'Ozone Watch' page](https://ozonewatch.gsfc.nasa.gov/facts/dobson_SH.html) for more information on atmospheric ozone.

```{r hide_src=TRUE}
ozone_vals <- seq(280, 420, by = 10)
```

### Effect of varying Ozone Total Column on PLC50 and ratio of PLC50:NLC50 for Anthracene

```{r hide_src=TRUE}
ozone_test_a <- multi_plc50(sites, 
                          pah = "Anthracene", 
                          varying = "o3_tc", 
                          vals = ozone_vals,
                          DOC = 5,
                          tauaer = 0.235)

ggplot(ozone_test_a, aes(x = o3_tc, y = plc50)) +
  geom_point() + 
  facet_wrap(vars(name), ncol = 1) +
  geom_point(
    data = filter(ozone_test_a, o3_tc == 300),
    colour = "red"
  ) + 
  geom_text_repel(
    data = filter(ozone_test_a, o3_tc == 300, name == "QUAMICHAN LAKE; CENTRE"),
    colour = "red", 
    label = "Recommended default = 300",
    nudge_x = 5, 
    nudge_y = 0.03
  ) +
  labs(
    title = "PLC50 for Anthracene at 3 sites in B.C. for a range of Ozone Total Column", 
    x = "Ozone Total Column",
    y = "PLC50 (Î¼g/L)"
  )

ggplot(ozone_test_a, aes(x = o3_tc, y = plc_nlc_ratio)) +
  geom_point() + 
  facet_wrap(vars(name), ncol = 1) +
  geom_point(
    data = filter(ozone_test_a, o3_tc == 300),
    colour = "red"
  ) + 
  geom_text_repel(
    data = filter(ozone_test_a, o3_tc == 300, name == "QUAMICHAN LAKE; CENTRE"),
    colour = "red", 
    label = "Recommended default = 300",
    nudge_x = 5, 
    nudge_y = 0.0005
  ) +
  labs(
    title = "Ratio of PLC50:NLC50 for Anthracene at 3 sites in B.C. for a range of\nOzone Total Column", 
    x = "Ozone Total Column",
    y = "Ratio of PLC50:NLC50"
  )
```

### Effect of varying Ozone Total Column on PLC50 and ratio of PLC50:NLC50 for Benzo[a]pyrene

```{r hide_src=TRUE}
ozone_test_b <- multi_plc50(sites, 
                          pah = "Benzo[a]pyrene", 
                          varying = "o3_tc", 
                          vals = ozone_vals,
                          DOC = 5,
                          tauaer = 0.235)

ggplot(ozone_test_b, aes(x = o3_tc, y = plc50)) +
  geom_point() + 
  facet_wrap(vars(name), ncol = 1) +
  geom_point(
    data = filter(ozone_test_b, o3_tc == 300),
    colour = "red"
  ) + 
  geom_text_repel(
    data = filter(ozone_test_b, o3_tc == 300, name == "QUAMICHAN LAKE; CENTRE"),
    colour = "red", 
    label = "Recommended default = 300",
    nudge_x = 5, 
    nudge_y = 0.0005
  ) +
  labs(
    title = "PLC50 for Benzo[a]pyrene at 3 sites in B.C. for a range of Ozone Total Column", 
    x = "Ozone Total Column",
    y = "PLC50 (Î¼g/L)"
  )

ggplot(ozone_test_b, aes(x = o3_tc, y = plc_nlc_ratio)) +
  geom_point() + 
  facet_wrap(vars(name), ncol = 1) +
  geom_point(
    data = filter(ozone_test_b, o3_tc == 300),
    colour = "red"
  ) + 
  geom_text_repel(
    data = filter(ozone_test_b, o3_tc == 300, name == "QUAMICHAN LAKE; CENTRE"),
    colour = "red", 
    label = "Recommended default = 300",
    nudge_x = 5, 
    nudge_y = 0.0003
  ) +
  labs(
    title = "Ratio of PLC50:NLC50 for Benzo[a]pyrene at 3 sites in B.C. for a range of\nOzone Total Column", 
    x = "Ozone Total Column",
    y = "Ratio of PLC50:NLC50"
  )
```

## Aerosol Optical Depth (`tauaer`)

The pahwq package currently uses average monthly aerosol optical depth data from 
2002 to 2023, obtained from MODIS/Aqua satellite data. Based on latitude and 
longitude and month, the AOD value is looked up and supplied to the TUV model. 
This default behaviour can be overridden by supplying a value to the `tauaer` 
argument in the `set_tuv_aq_params` function. In the absence of
climatological data, the recommended default value is 0.235 (@aris2023).

To test the sensitivity of PLC50 to AOD, a range of values is tested from 0.1 
(clear skies with maximum visibility) to 1.0 (very hazy skies). See the 
[NASA Earth Observatory page on aerosols](https://earthobservatory.nasa.gov/global-maps/MODAL2_M_AER_OD).

```{r hide_src=TRUE}
aod_vals <- c(seq(0.1, 1.0, length.out = 10), 0.235)
```

### Effect of Varying Aerosol Optical Depth (AOD) on PLC50 and ratio of PLC50:NLC50 for Anthracene

```{r hide_src=TRUE}
aod_test_a <- multi_plc50(sites, 
                        pah = "Anthracene", 
                        varying = "tauaer", 
                        vals = aod_vals,
                        DOC = 5,
                        o3_tc = 300)

ggplot(aod_test_a, aes(x = tauaer, y = plc50)) +
  geom_point() + 
  facet_wrap(vars(name), ncol = 1) +
  geom_point(
    data = filter(aod_test_a, tauaer == 0.235),
    colour = "red"
  ) + 
  geom_text_repel(
    data = filter(aod_test_a, tauaer == 0.235, name == "QUAMICHAN LAKE; CENTRE"),
    colour = "red", 
    label = "Recommended default = 0.235",
    nudge_y = 0.1
  ) +
  labs(
    title = "PLC50 for Anthracene at 3 sites in B.C. for a range of Aeorosol Optical Depth", 
    x = "Aeorosol Optical Depth",
    y = "PLC50 (Î¼g/L)"
  )

ggplot(aod_test_a, aes(x = tauaer, y = plc_nlc_ratio)) +
  geom_point() + 
  facet_wrap(vars(name), ncol = 1) +
  geom_point(
    data = filter(aod_test_a, tauaer == 0.235),
    colour = "red"
  ) + 
  geom_text_repel(
    data = filter(aod_test_a, tauaer == 0.235, name == "QUAMICHAN LAKE; CENTRE"),
    colour = "red", 
    label = "Recommended default = 0.235",
    nudge_y = 0.002
  ) +
  labs(
    title = "Ratio of PLC50:NLC50 for Anthracene at 3 sites in B.C. for a range of\nAeorosol Optical Depth", 
    x = "Aeorosol Optical Depth",
    y = "Ratio of PLC50:NLC50"
  )
```

### Effect of varying Aerosol Optical Depth (AOD) on PLC50 and ratio of PLC50:NLC50 for Benzo[a]pyrene

```{r hide_src=TRUE}
aod_test_b <- multi_plc50(sites, 
                        pah = "Benzo[a]pyrene", 
                        varying = "tauaer", 
                        vals = aod_vals,
                        DOC = 5,
                        o3_tc = 300)

ggplot(aod_test_b, aes(x = tauaer, y = plc50)) +
  geom_point() + 
  facet_wrap(vars(name), ncol = 1) +
  geom_point(
    data = filter(aod_test_b, tauaer == 0.235),
    colour = "red"
  ) + 
  geom_text_repel(
    data = filter(aod_test_b, tauaer == 0.235, name == "QUAMICHAN LAKE; CENTRE"),
    colour = "red", 
    label = "Recommended default = 0.235",
    nudge_y = 0.002
  ) +
  labs(
    title = "PLC50 for Benzo[a]pyrene at 3 sites in B.C. for a range of\nAeorosol Optical Depth", 
    x = "Aeorosol Optical Depth",
    y = "PLC50 (Î¼g/L)"
  )

ggplot(aod_test_b, aes(x = tauaer, y = plc_nlc_ratio)) +
  geom_point() + 
  facet_wrap(vars(name), ncol = 1) +
  geom_point(
    data = filter(aod_test_b, tauaer == 0.235),
    colour = "red"
  ) + 
  geom_text_repel(
    data = filter(aod_test_b, tauaer == 0.235, name == "QUAMICHAN LAKE; CENTRE"),
    colour = "red", 
    label = "Recommended default = 0.235",
    nudge_y = 0.001
  ) +
  labs(
    title = "Ratio of PLC50:NLC50 for Benzo[a]pyrene at 3 sites in B.C. for a range of\nAeorosol Optical Depth", 
    x = "Aeorosol Optical Depth",
    y = "Ratio of PLC50:NLC50"
  )
```

## Aerosol Single Scattering Albedo (`ssaaer`)

Aerosol Single Scattering Albedo is currently set as a default constant value of
0.99. This can be overridden by setting the `ssaaer` parameter of the 
`set_tuv_aq_params()` function to a different value. 

Here we test a range for values from 0.80 to 0.99, including the recommended 
default of 0.99 (@aris2023).

```{r hide_src=TRUE}
ssa_vals <- seq(0.8, 0.99, length.out = 10)
```

### Effect of varying Aerosol Single Scattering Albedo on PLC50 and ratio of PLC50:NLC50 for Anthracene

```{r hide_src=TRUE}
ssa_test_a <- multi_plc50(sites, 
                        pah = "Anthracene", 
                        varying = "ssaaer", 
                        vals = ssa_vals,
                        DOC = 5,
                        o3_tc = 300,
                        tauaer = 0.235)

ggplot(ssa_test_a, aes(x = ssaaer, y = plc50)) +
  geom_point() + 
  facet_wrap(vars(name), ncol = 1) +
  geom_point(
    data = filter(ssa_test_a, ssaaer == 0.99),
    colour = "red"
  ) + 
  geom_text_repel(
    data = filter(ssa_test_a, ssaaer == 0.99, name == "QUAMICHAN LAKE; CENTRE"),
    colour = "red", 
    label = "Recommended default = 0.99",
    nudge_y = .06
  ) +
  labs(
    title = "PLC50 for Anthracene at 3 sites in B.C. for a range of\nAeorosol Single Scattering Albedo", 
    x = "Aeorosol Single Scattering Albedo",
    y = "PLC50 (Î¼g/L)"
  )

ggplot(ssa_test_a, aes(x = ssaaer, y = plc_nlc_ratio)) +
  geom_point() + 
  facet_wrap(vars(name), ncol = 1) +
  geom_point(
    data = filter(ssa_test_a, ssaaer == 0.99),
    colour = "red"
  ) + 
  geom_text_repel(
    data = filter(ssa_test_a, ssaaer == 0.99, name == "QUAMICHAN LAKE; CENTRE"),
    colour = "red", 
    label = "Recommended default = 0.99",
    nudge_y = .001
  ) +
  labs(
    title = "Ratio of PLC50:NLC50 for Anthracene at 3 sites in B.C. for a range of\nAeorosol Single Scattering Albedo", 
    x = "Aeorosol Single Scattering Albedo",
    y = "Ratio of PLC50:NLC50"
  )
```

### Effect of varying Aerosol Single Scattering Albedo on PLC50 and ratio of PLC50:NLC50 for Benzo[a]pyrene

```{r hide_src=TRUE}
ssa_test_b <- multi_plc50(sites, 
                        pah = "Benzo[a]pyrene", 
                        varying = "ssaaer", 
                        vals = ssa_vals,
                        DOC = 5,
                        o3_tc = 300,
                        tauaer = 0.235)

ggplot(ssa_test_b, aes(x = ssaaer, y = plc50)) +
  geom_point() + 
  facet_wrap(vars(name), ncol = 1) +
  geom_point(
    data = filter(ssa_test_b, ssaaer == 0.99),
    colour = "red"
  ) + 
  geom_text_repel(
    data = filter(ssa_test_b, ssaaer == 0.99, name == "QUAMICHAN LAKE; CENTRE"),
    colour = "red", 
    label = "Recommended default = 0.99",
    nudge_y = .001
  ) +
  labs(
    title = "PLC50 for Benzo[a]pyrene at 3 sites in B.C. for a range of\nAeorosol Single Scattering Albedo", 
    x = "Aeorosol Single Scattering Albedo",
    y = "PLC50 (Î¼g/L)"
  )

ggplot(ssa_test_b, aes(x = ssaaer, y = plc_nlc_ratio)) +
  geom_point() + 
  facet_wrap(vars(name), ncol = 1) +
  geom_point(
    data = filter(ssa_test_b, ssaaer == 0.99),
    colour = "red"
  ) + 
  geom_text_repel(
    data = filter(ssa_test_b, ssaaer == 0.99, name == "QUAMICHAN LAKE; CENTRE"),
    colour = "red", 
    label = "Recommended default = 0.99",
    nudge_y = .0005
  ) +
  labs(
    title = "Ratio of PLC50:NLC50 for Benzo[a]pyrene at 3 sites in B.C. for a range of\nAeorosol Single Scattering Albedo", 
    x = "Aeorosol Single Scattering Albedo",
    y = "Ratio of PLC50:NLC50"
  )
```

## Latitude (`lat`)

Latitude has a strong effect on the angle of the sun, which will in turn affect
the light penetration through water, thus it is important to investigate the 
sensitivity of the calculation of PLC50 to variation in latitude.

For this analysis, we will maintain the longitudes associated with each site, 
and make "mock" sites by creating a range of latitudes from 48 to 72 degrees N 
and pair those with the longitudes of the original sites.

```{r}
lat_vals <- seq(48, 72, by = 2)
```

### Effect of varying Latitude on PLC50 and ratio of PLC50:NLC50 for Anthracene

```{r hide_src=TRUE}
lat_test_a <- multi_plc50(sites, 
                        pah = "Anthracene", 
                        varying = "lat", 
                        vals = lat_vals,
                        DOC = 5,
                        o3_tc = 300,
                        tauaer = 0.235)

ggplot(lat_test_a, aes(x = lat, y = plc50)) +
  geom_point() + 
  facet_wrap(vars(lon), ncol = 1, labeller = as_labeller(\(x) paste("Longitude: ", x))) +
  labs(
    title = "PLC50 for Anthracene at 3 sites in B.C. for a range of latitudes", 
    x = "Latitude",
    y = "PLC50 (Î¼g/L)"
  )

ggplot(lat_test_a, aes(x = lat, y = plc_nlc_ratio)) +
  geom_point() + 
  facet_wrap(vars(lon), ncol = 1, labeller = as_labeller(\(x) paste("Longitude: ", x))) +
  labs(
    title = "Ratio of PLC50:NLC50 for Anthracene at 3 sites in B.C. for a range of latitudes",
    x = "Latitude",
    y = "Ratio of PLC50:NLC50"
  )
```

### Effect of varying Latitude on PLC50 and ratio of PLC50:NLC50 for Benzo[a]pyrene

```{r hide_src=TRUE}
lat_test_b <- multi_plc50(sites, 
                        pah = "Benzo[a]pyrene", 
                        varying = "lat", 
                        vals = lat_vals,
                        DOC = 5,
                        o3_tc = 300,
                        tauaer = 0.235)

ggplot(lat_test_b, aes(x = lat, y = plc50)) +
  geom_point() + 
  facet_wrap(vars(lon), ncol = 1, labeller = as_labeller(\(x) paste("Longitude: ", x))) +
  labs(
    title = "PLC50 for Benzo[a]pyrene at 3 sites in B.C. for a range of latitudes", 
    x = "Latitude",
    y = "PLC50 (Î¼g/L)"
  )

ggplot(lat_test_b, aes(x = lat, y = plc_nlc_ratio)) +
  geom_point() + 
  facet_wrap(vars(lon), ncol = 1, labeller = as_labeller(\(x) paste("Longitude: ", x))) +
  labs(
    title = "Ratio of PLC50:NLC50 for Benzo[a]pyrene at 3 sites in B.C. for a range of\nlatitudes",
    x = "Latitude",
    y = "Ratio of PLC50:NLC50"
  )
```

## Elevation (`elev_km`)

We can test the effect of elevation on PLC50 by varying elevation from sea level 
to 1000m (1km) above sea level. We will keep the latitude and longitude associated 
with each site.

```{r}
elevation_vals <- seq(0, 1, by = 0.1)
```

### Effect of varying Elevation on PLC50 and ratio of PLC50:NLC50 for Anthracene

```{r hide_src=TRUE}
elev_test_a <- multi_plc50(sites, 
                         pah = "Anthracene", 
                         varying = "elev_km", 
                         vals = elevation_vals,
                         DOC = 5,
                         o3_tc = 300,
                         tauaer = 0.235)

ggplot(elev_test_a, aes(x = elev_km, y = plc50)) +
  geom_point() + 
  facet_wrap(vars(name), ncol = 1) +
  labs(
    title = "PLC50 for Anthracene at 3 sites in B.C. for a range of elevations", 
    x = "Elevation (km)",
    y = "PLC50 (Î¼g/L)"
  )

ggplot(elev_test_a, aes(x = elev_km, y = plc_nlc_ratio)) +
  geom_point() + 
  facet_wrap(vars(name), ncol = 1) +
  labs(
    title = "Ratio of PLC50:NLC50 for Anthracene at 3 sites in B.C. for a range of\nelevations",
    x = "Elevation (km)",
    y = "Ratio of PLC50:NLC50"
  )
```

### Effect of varying Elevation on PLC50 and ratio of PLC50:NLC50 for Benzo[a]pyrene

```{r hide_src=TRUE}
elev_test_b <- multi_plc50(sites, 
                         pah = "Benzo[a]pyrene", 
                         varying = "elev_km", 
                         vals = elevation_vals,
                         DOC = 5,
                         o3_tc = 300,
                         tauaer = 0.235)

ggplot(elev_test_b, aes(x = elev_km, y = plc50)) +
  geom_point() + 
  facet_wrap(vars(name), ncol = 1) +
  labs(
    title = "PLC50 for Benzo[a]pyrene at 3 sites in B.C. for a range of elevations", 
    x = "Elevation (km)",
    y = "PLC50 (Î¼g/L)"
  )

ggplot(elev_test_b, aes(x = elev_km, y = plc_nlc_ratio)) +
  geom_point() + 
  facet_wrap(vars(name), ncol = 1) +
  labs(
    title = "Ratio of PLC50:NLC50 for Benzo[a]pyrene at 3 sites in B.C. for a range of\nelevations",
    x = "Elevation (km)",
    y = "Ratio of PLC50:NLC50"
  )
```

## Water depth (`depth_m`)

The depth at which the photoxicity of a PAH is determined 
should be a conservative one - i.e., a depth at which sensitive species/life stages exist and thus would be exposed. The default is set at 0.25m.

We can test the effect of depth on PLC50 by varying
elevation from water level to 1m below the surface:

```{r}
depth_vals <- c(seq(0, 1, by = 0.1), 0.25)
```

### Effect of varying Water Depth on PLC50 and ratio of PLC50:NLC50 for Anthracene

```{r hide_src=TRUE}
depth_test_a <- multi_plc50(sites, 
                          pah = "Anthracene", 
                          varying = "depth_m", 
                          vals = depth_vals,
                          DOC = 5,
                          o3_tc = 300,
                          tauaer = 0.235)

ggplot(depth_test_a, aes(x = depth_m, y = plc50)) +
  geom_point() + 
  geom_point(
    data = filter(depth_test_a, depth_m == 0.25),
    colour = "red"
  ) + 
  geom_text_repel(
    data = filter(depth_test_a, depth_m == 0.25, name == "QUAMICHAN LAKE; CENTRE"),
    colour = "red", 
    label = "Recommended default = 0.25m",
    nudge_x = -0.05,
    nudge_y = 5
  ) +
  facet_wrap(vars(name), ncol = 1) +
  labs(
    title = "PLC50 for Anthracene at 3 sites in B.C. for a range of water depths", 
    x = "Water Depth (m)",
    y = "PLC50 (Î¼g/L)"
  )

ggplot(depth_test_a, aes(x = depth_m, y = plc_nlc_ratio)) +
  geom_point() + 
  geom_point(
    data = filter(depth_test_a, depth_m == 0.25),
    colour = "red"
  ) + 
  geom_text_repel(
    data = filter(depth_test_a, depth_m == 0.25, name == "QUAMICHAN LAKE; CENTRE"),
    colour = "red", 
    label = "Recommended default = 0.25m",
    nudge_x = -0.05,
    nudge_y = 0.1
  ) +
  facet_wrap(vars(name), ncol = 1) +
  labs(
    title = "Ratio of PLC50:NLC50 for Anthracene at 3 sites in B.C. for a range of\nwater depths",
    x = "Water Depth (m)",
    y = "Ratio of PLC50:NLC50"
  )
```

### Effect of varying Water Depth on PLC50 and ratio of PLC50:NLC50 for Benzo[a]pyrene

```{r hide_src=TRUE}
depth_test_b <- multi_plc50(sites, 
                          pah = "Benzo[a]pyrene", 
                          varying = "depth_m", 
                          vals = depth_vals,
                          DOC = 5,
                          o3_tc = 300,
                          tauaer = 0.235)

ggplot(depth_test_b, aes(x = depth_m, y = plc50)) +
  geom_point() + 
  geom_point(
    data = filter(depth_test_b, depth_m == 0.25),
    colour = "red"
  ) + 
  geom_text_repel(
    data = filter(depth_test_b, depth_m == 0.25, name == "QUAMICHAN LAKE; CENTRE"),
    colour = "red", 
    label = "Recommended default = 0.25m",
    nudge_x = -0.05,
    nudge_y = 0.1
  ) +
  facet_wrap(vars(name), ncol = 1) +
  labs(
    title = "PLC50 for Benzo[a]pyrene at 3 sites in B.C. for a range of water depths", 
    x = "Water Depth (m)",
    y = "PLC50 (Î¼g/L)"
  )

ggplot(depth_test_b, aes(x = depth_m, y = plc_nlc_ratio)) +
  geom_point() + 
  geom_point(
    data = filter(depth_test_b, depth_m == 0.25),
    colour = "red"
  ) + 
  geom_text_repel(
    data = filter(depth_test_b, depth_m == 0.25, name == "QUAMICHAN LAKE; CENTRE"),
    colour = "red", 
    label = "Recommended default = 0.25m",
    nudge_x = -0.05,
    nudge_y = 0.05
  ) +
  facet_wrap(vars(name), ncol = 1) +
  labs(
    title = "Ratio of PLC50:NLC50 for Benzo[a]pyrene at 3 sites in B.C. for a range of\nwater depths",
    x = "Water Depth (m)",
    y = "Ratio of PLC50:NLC50"
  )
```


## Radiative transfer scheme (`nstr`)

Number of streams for radiative transfer calculations: 

* If `nstr` < 2, uses 2-stream delta-Eddington (faster)
* if `nstr` >= 2, uses n-stream discrete ordinates (more accurate: must be even number, maximum = 32)

```{r, hide_src=TRUE, cache=TRUE, fig.height=6}
out <- multi_plc50(sites, 
          pah = "Anthracene", 
          varying = "nstr", 
          vals = c(-2, 4), 
          DOC = 5,
          o3_tc = 300,
          tauaer = 0.235) |> 
  mutate(nstr = factor(nstr))

ggplot(out, aes(x = name, y = plc50, colour = nstr)) + 
  geom_point() + 
  labs(
    title = "Calculated PLC50 using two different TUV methods for radiative transfer", 
    x = "Site"
  ) + 
  theme(axis.text.x = element_text(angle = 355))

ggplot(out, aes(x = name, y = timing, colour = nstr)) + 
  geom_point() + 
  labs(
    title = "Execution time of TUV model runs with different methods for radiative transfer", 
    x = "Site",
    y = "Execution Time (s)"
  ) + 
  theme(axis.text.x = element_text(angle = 355))
```

We can see that there are small differences in the calculated PLC50 when using different values of `nstr`, with the values when calculated using `nstr = -2` being on average 
`r round(out |> group_by(name) |> summarize(diff = (plc50[nstr == 4] - plc50[nstr == -2]) / plc50[nstr == 4]) |> pull(diff) |> mean() * 100, 2)`% lower than when using `nstr = 4`. Using `nstr = 4` however, is about `r round(out |> group_by(name) |> summarize(diff = (timing[nstr == 4] / timing[nstr == -2])) |> pull(diff) |> mean())` times slower than using `nstr = -2`. In most cases, unless high precision is required, it is likely that using the much faster `nstr = -2` should be the default.

## Methylated Polycyclic Aromatic Hydrocarbons

In this section we will look at the relative photoxicity of methylated PAHs as
compared to their parent compounds.

For these comparisons we will limit it to just one site in the Okanagan.

```{r}
okanagan <- filter(sites, emsid == "0500236")
```

### Napthalene derivatives:

Before we calculate the photoxicity of various methylated Naphthalene compounds, 
we can examine their molar absorption rates across the spectrum of interest:

```{r, hide_src=TRUE, fig.height=6}
pahwq:::molar_absorption |> 
  filter(grepl("naphthalene", chemical)) |> 
  mutate(
    chemical = factor(chemical, levels = c("naphthalene", "1-methylnaphthalene", "2,6-dimethylnaphthalene", "1,6,7-trimethylnaphthalene"))
  ) |> 
  ggplot(aes(x = wavelength, y = molar_absorption, colour = chemical)) + 
  geom_point() + 
  labs(
    title = "Molar absorption of Naphthalene and methylated derivatives", 
    y = "Molar absorption (L/mol/cm)"
  )
```

We can see in general that more highly methylated compounds have greater absorption, though it is not consistent across the spectrum.

We can also examine the properties and narcotic toxicity (NLC50) of Naphthalene
and its methylated variants:

```{r, hide_src=TRUE, fig.height=6}
pahwq:::nlc50_lookup |> 
  filter(grepl("naphthalene$", chemical)) |> 
  mutate(
    nlc50 = vapply(chemical, nlc50, FUN.VALUE = 1), 
    chemical = factor(chemical, levels = c("naphthalene", "1-methylnaphthalene", "2,6-dimethylnaphthalene", "1,6,7-trimethylnaphthalene"))
  ) |> 
  ggplot(aes(x = log_kow, y = nlc50, colour = chemical)) + 
  geom_point() + 
  labs(
    title = "NLC50 of Naphthalene and its methylated derivatives",
    y = "NLC50 (ug/L)"
  )
```

We can see here that 1,6,7-Trimethylnaphthalene has the highest log(KoW) and lowest NLC50.

Finally, we can compare the phototoxicity (PLC50) of Naphthalene and its methylated
derivatives at the Okanagan site, at the two values of $k_{d,305}$ (1 and 150).

```{r, hide_src=TRUE}
naphthalene <- multi_plc50(
  okanagan,
  site = "name",
  pah = c("Naphthalene", "1-Methylnaphthalene", "2,6-Dimethylnaphthalene", "1,6,7-Trimethylnaphthalene"),
  varying = "Kd_ref",
  vals = c(1, 150),
  depth_m = 0.25
) |> 
  mutate(PAH = factor(PAH, levels = c("Naphthalene", "1-Methylnaphthalene", "2,6-Dimethylnaphthalene", "1,6,7-Trimethylnaphthalene")))

ggplot(naphthalene, aes(x = PAH, y = plc50, colour = PAH)) + 
  geom_point() + 
  facet_wrap(vars(Kd_ref), ncol = 1, scales = "free_y", 
             labeller = as_labeller(function(x) paste0("Kd(305) = ", x))) + 
  theme(axis.text.x = element_text(angle = 355)) + 
  scale_color_discrete(guide = "none") + 
  labs(
    title = "Comparison of PLC50 of Naphthalene and its methylated derivatives",
    y = "PLC50 (ug/L)"
  )
```

Here we can see that the methylated compounds are all more phototoxic than the 
parent Naphthalene, with the most methylated variant being the most toxic (lowest PLC50).

### Phenanthrene derivatives

9-methylphenanthrene vs phenanthrene

### Anthracene derivatives

9-methylanthracene vs anthracene

### Pyrene derivatives

1-methylpyrene vs pyrene

### Chrysene derivatives

1-methylchrysene vs chrysene

## References
